---
title: "RPA_Reweighter"
format: html
editor: visual
project: execute-dir: project
---

## RPA Reweighter

This set of scripts runs the household reweighter to produce reweighted PUMS data that satisfies household targets set by the user.

# 1 - Set Scenario Name, knitr path, and root directory, load packages

```{r, setup, include=FALSE}

pacman::p_load(
  tidyverse,
  tidycensus,
  mapcdatakeys,
  janitor,
  readxl,
  rjson,
  stringr,
  data.table,
  here
)
# Set Scenario name

scen <- 'Scenario_3'
# scen <- 'Scenario_2'
# scen <- 'Scenario_3'
# scen <- 'Scenario_4'

# Set ACS / PUMS vintage
vintage <-  2021

# Set headship rate scenario
#hdshp <- 'acs1721'
hdshp <- 'scen2'

# Set MAPC at 101 or 97 towns
reb <- 'MAPC101'
# reb <- 'MAPC97'

# List of RPA names
mpos <- unlist(c(mapcdatakeys::all_muni_data_keys |> select(mpo) |> unique()))

# Root Directory
root <- 'S:/Network Shares/K Drive/DataServices/Projects/'
# root <- 'K:/DataServices/Projects/'

umdi.path <- paste0(root, 'Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Working/Regional_Control_Totals/UMDI/', scen, '/')

ctrltotal.path <- paste0(root, 'Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Working/Regional_Control_Totals/', scen, '/UPDATE/')

#ctrltotal.path <- paste0(root, 'Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Working/Regional_Control_Totals/', scen, '/')

workpath <- paste0(root, 'Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Working/Reweighter/')

input_path <- paste0(workpath, scen, '/Input_Files/') 
output_path <- paste0(workpath, scen, '/Output_Files/') 

# Subfolders for RPA inputs and outputs
dir.create(paste0(workpath, scen, '/Input_Files/'), showWarnings = F)
dir.create(paste0(workpath, scen, '/Output_Files/'), showWarnings = F)
  for (m in mpos) {
    dir.create(paste0(workpath, scen, '/Input_Files/', m), showWarnings = F)
    dir.create(paste0(workpath, scen, '/Output_Files/', m), showWarnings = F)
  }

#Set knitr path
#knitr::opts_knit$set(root.dir = workpath)

## turn off scientific notation
options(scipen = 999)
#Set random seed
set.seed(351)
# Helper function
sun <- function(x){sort(unique(x))}
lun <- function(x){length(unique(x))}

```

# 2 - Subset PUMS data for each RPA

Formatted PUMS data is created by the *create_regional_control_totals.qmd* in section 6.2

# 2.1 - Scenario 1

```{r}
if(scen=='Scenario_1') {
  pums_data <- fread(paste0(workpath, 'PUMS_data/PUMS', vintage, '_formatted.csv')) |> filter(TYPEHUGQ == "1") |>
    mutate(AgeCat = ageCAT6)
  
  mapc.pumas <- pums_data[rpa_acr == 'MAPC', unique(PUMA)]
  swm.pumas <- pums_data[rpa_acr != 'MAPC', .(rpa_acr, PUMA)] |>
    unique()
  # MAPC 101 subset
  fwrite(
    pums_data[PUMA %in% mapc.pumas],
    paste0(
      workpath,
      scen,
      '/Input_Files/MAPC/PUMS2021_',
      scen,
      '_MAPC.csv'
    )
  )
  message(paste0('MAPC 101 PUMS export complete ', yr))
  # SWM join PUMA and export subsets
  for (m in unique(swm.pumas$rpa_acr)) {
    pums_data |>
      filter(rpa_acr == m) |>
      fwrite(
        paste0(
          workpath,
          scen,
          '/Input_Files/',
          m,
          '/PUMS2021_',
          scen,
          '_',
          m,
          '.csv'
        )
      )
    message(paste0(m, ' PUMS export complete ', yr))
  }
}
```

# 2.2 - Scenario 2/3/4

```{r}
if(scen %in% c('Scenario_2', 'Scenario_3', 'Scenario_4')) {
  
  # We keep the base ACS 17-21 headship rates and PUMS subset from Scenario 1 for the year 2025, but applied to the Scenario 2/3/4 populations from UMDI. Adjusted headship rates and PUMS subsets are used only for year 2030 and after
  for (yr in 2025){
    pums_data <- fread(paste0(workpath, 'PUMS_data/PUMS', vintage, '_formatted.csv')) |> filter(TYPEHUGQ == "1") |>
    mutate(AgeCat = ageCAT6)
  dir.create(paste0(workpath, scen, '/Input_Files/'), showWarnings = F)
  dir.create(paste0(workpath, scen, '/Output_Files/'), showWarnings = F)
  for (m in mpos) {
    dir.create(paste0(workpath, scen, '/Input_Files/', m), showWarnings = F)
    dir.create(paste0(workpath, scen, '/Output_Files/', m), showWarnings = F)
  }
  mapc.pumas <- pums_data[rpa_acr == 'MAPC', unique(PUMA)]
  swm.pumas <- pums_data[rpa_acr != 'MAPC', .(rpa_acr, PUMA)] |>
    unique()
  # MAPC 101 subset
  fwrite(
    pums_data[PUMA %in% mapc.pumas],
    paste0(
      workpath,
      scen,
      '/Input_Files/MAPC/PUMS2021_',
      scen,
      '_MAPC_',
      yr,
      '.csv'
    )
  )
  message(paste0('MAPC 101 PUMS export complete ', yr))
  # SWM join PUMA and export subsets
  for (m in unique(swm.pumas$rpa_acr)) {
    pums_data |>
      filter(rpa_acr == m) |>
      fwrite(
        paste0(
          workpath,
          scen,
          '/Input_Files/',
          m,
          '/PUMS2021_',
          scen,
          '_',
          m,
          '_',
          yr,
          '.csv'
        )
      )
    message(paste0(m, ' PUMS export complete ', yr))
  }
  }
  
  for (yr in seq(2030, 2050, 5)) {
    x2 <- fread(paste0(workpath, 'PUMS_data/PUMA10_RPA_crosswalk.csv')) |> 
      mutate(PUMA = as.numeric(PUMA))
    pums_data <- fread(paste0(workpath, 'PUMS_data/PUMS2021_Scenario_2_', yr, '.csv')) |>
      left_join(x2, by = c('PUMA')) |>
      filter(TYPEHUGQ == "1") |>
      mutate(AgeCat = ageCAT6)
    pums_data[!is.na(SPORDER.adj), SPORDER := SPORDER.adj]
    pums_data[!is.na(HHtype.adj), HHtype := HHtype.adj]
    pums_data[!is.na(SERIALNO.adj), SERIALNO := SERIALNO.adj]

    mapc.pumas <- pums_data[rpa_acr == 'MAPC', unique(PUMA)]
    swm.pumas <- pums_data[rpa_acr != 'MAPC', .(rpa_acr, PUMA)] |>
      unique()
    
    # MAPC 101 subset
    fwrite(
      pums_data[PUMA %in% mapc.pumas],
      paste0(
        workpath,
        scen,
        '/Input_Files/MAPC/PUMS2021_',
        scen,
        '_MAPC_',
        yr,
        '.csv'
      )
    )
    message(paste0('MAPC 101 PUMS export complete ', yr))
    
    # SWM join PUMA and export subsets
    for (m in unique(swm.pumas$rpa_acr)) {
      pums_data |>
        filter(rpa_acr == m) |>
        fwrite(
          paste0(
            workpath,
            scen,
            '/Input_Files/',
            m,
            '/PUMS2021_',
            scen,
            '_',
            m,
            '_',
            yr,
            '.csv'
          )
        )
      message(paste0(m, ' PUMS export complete ', yr))
    }
  }
}
```

# 3 - Create RPA config JSON files for reweighter configuration

```{r}

# Template config file
infile1 <- paste0(workpath, 'Templates/reweighting_config_2021.json')

infile2 <- paste0(workpath, 'Templates/reweighting_config_2020baseyear.json')

for (m in mpos){
  outfile1 <- paste0(workpath, scen, '/Input_Files/', m, '/reweighting_config_2021_', m, '.json')
  outfile2 <- paste0(workpath, scen, '/Input_Files/', m, '/reweighting_config_2020baseyear_', m, '.json')
  cfig1 = readLines(infile1)
  cfig1 = gsub('PUMS2021', paste0('PUMS2021_', m), cfig1)
  writeLines(cfig1, outfile1)
  cfig2 = readLines(infile2)
  cfig2 = gsub('PUMS2021', paste0('PUMS2021_', m), cfig2)
  writeLines(cfig2, outfile2)
}

```

# 4 - Configure, set targets and run reweighter for each RPA and year

```{r}

#



source('algo_hh.R')

# Set Targets and run reweighter loop

# Different configuration for 2020 base year
if (scen == 'Scenario_1') {
  for (yr in c(2020)) {
    #for (mid in c('MAPC')) {
    for (mid in mpos) {
      # Set config files
      inpath <- paste0(workpath, scen, '/Input_Files/', mid, '/')
      
      source(here('scripts/reweighter/read_config_RPA_2020baseyear.R'))
      message(paste0(mid, ' configuration complete'))
      
      outpath <- paste0(workpath, scen, '/Output_Files/', mid, '/')
      age.target <- fread(
        paste0(
          ctrltotal.path,
          'age.targetfile.',
          reb,
          '.',
          scen,
          '.hdshp.',
          hdshp,
          '.y',
          yr,
          '.csv'
        )
      ) |>
        filter(RPA == mid) |>
        select(-RPA) |> setDT()
      setorder(age.target, ageCAT4)
      fwrite(age.target, paste0(inpath, '/age.csv'))
      
      size.target <- fread(
        paste0(
          ctrltotal.path,
          'size.targetfile.',
          reb,
          '.',
          scen,
          '.hdshp.',
          hdshp,
          '.y',
          yr,
          '.csv'
        )
      ) |>
        filter(RPA == mid) |>
        select(-RPA) |> setDT()
      setorder(size.target, size)
      fwrite(size.target, paste0(inpath, '/size.csv'))
      
      type.target <- fread(
        paste0(
          ctrltotal.path,
          'type.targetfile.',
          reb,
          '.',
          scen,
          '.hdshp.',
          hdshp,
          '.y',
          yr,
          '.csv'
        )
      ) |>
        filter(RPA == mid) |>
        select(-RPA) |> setDT()
      setorder(type.target, HHtype)
      fwrite(type.target, paste0(inpath, '/type.csv'))
      
      message(paste0(mid, ' targets set'))
      
      pums <- fread(paste0(inpath, 'PUMS', vintage, '_', mid, '.csv'))
      
      source(here('scripts/reweighter/Manage_Scenario_RPA.R'))
      message(paste0(mid, ' reweighter complete'))
      
      mess <- data.table(txt = paste0(mid, ' reweighter complete'))
      fwrite(mess, paste0(outpath, mid, '_', yr, '_complete.txt'))
    }
  }
  for (yr in c(2025, 2030, 2035, 2040, 2045, 2050)) {
    for (mid in mpos) {
      # Set config
      inpath <- paste0(workpath, scen, '/Input_Files/', mid, '/')
      source(here('scripts/reweighter/read_config_RPA.R'))
      message(paste0(mid, ' configuration complete'))
      
      # Set targets
      outpath <- paste0(workpath, scen, '/Output_Files/', mid, '/')
      dir.create(outpath, showWarnings = F)
      template <- fread(paste0(workpath, 'Templates/AgeHHder_HHtype_template.csv')) |>
        select(-BASELINE, -TARGET)
      target <- fread(
        paste0(
          ctrltotal.path,
          'hhtype.targetfile.',
          reb,
          '.',
          scen,
          '.hdshp.',
          hdshp,
          '.y',
          yr,
          '.csv'
        )
      ) |>
        filter(RPA == mid) |>
        select(-RPA) |> right_join(template, by = c('AgeCat', 'HHtype')) |> setDT()
      target[AgeCat %in% c('x == 1', 'x == 2', 'x == 3'), `:=`(BASELINE =
                                                                 0,
                                                               INTER = 0,
                                                               TARGET = 0)]
      target[is.na(BASELINE), `:=`(BASELINE = 0,
                                   INTER = 0,
                                   TARGET = 0)]
      setorder(target, AgeCat, HHtype)
      fwrite(target, paste0(inpath, '/AgeHHder_HHtype.csv'))
      
      template2 <- fread(paste0(workpath, 'Templates/age_tenure_template.csv'))
      target2 <- fread(
        paste0(
          ctrltotal.path,
          "tenure.targetfile.",
          reb,
          ".",
          scen,
          '.hdshp.',
          hdshp,
          ".y",
          yr,
          ".csv"
        )
      ) |>
        filter(RPA == mid) |>
        select(-RPA) |>
        right_join(template2, by = c('AgeCat', 'tenure')) |>
        setDT()
      
      target2[is.na(BASELINE), `:=`(BASELINE = 0,
                                    INTER = 0,
                                    TARGET = 0)]
      setorder(target2, AgeCat, tenure)
      fwrite(target2, paste0(inpath, '/age_tenure.csv'))
      
      message(paste0(mid, ' targets set'))
      
      pums <- fread(paste0(inpath, 'PUMS', vintage, '_', mid, '.csv'))
      if (scen == 'Scenario_2') {
        pums[, SPORDER := SPORDER.adj][, SERIALNO := SERIALNO.adj]
      }
      
      source(here('scripts/reweighter/Manage_Scenario_RPA.R'))
      message(paste0(mid, ' reweighter complete'))
      
      mess <- data.table(txt = paste0(mid, ' reweighter complete'))
      fwrite(mess, paste0(outpath, mid, '_', yr, '_complete.txt'))
    }
  }
}

if (scen %in% c('Scenario_2', 'Scenario_3', 'Scenario_4')) {
  for (yr in c(2025, 2030, 2035, 2040, 2045, 2050)) {
    for (mid in mpos) {
      infile1 <- paste0(workpath, 'Templates/reweighting_config_2021.json')
      outfile1 <- paste0(
        workpath,
        scen,
        '/Input_Files/',
        mid,
        '/reweighting_config_2021_',
        mid,
        '.json'
      )
      cfig1 = readLines(infile1)
      cfig1 = gsub('PUMS2021.csv',
                   paste0('PUMS2021_', scen, '_', mid, '_', yr, '.csv'),
                   cfig1)
      writeLines(cfig1, outfile1)
      # Set config
      inpath <- paste0(workpath, scen, '/Input_Files/', mid, '/')
      source(here('scripts/reweighter/read_config_RPA.R'))
      message(paste0(mid, ' configuration complete'))
      
      # Set targets
      outpath <- paste0(workpath, scen, '/Output_Files/', mid, '/')
      dir.create(outpath, showWarnings = F)
      template <- fread(paste0(workpath, 'Templates/AgeHHder_HHtype_template.csv')) |>
        select(-BASELINE, -TARGET)
      target <- fread(
        paste0(
          ctrltotal.path,
          'hhtype.targetfile.',
          reb,
          '.',
          scen,
          '.hdshp.',
          hdshp,
          '.y',
          yr,
          '.csv'
        )
      ) |>
        filter(RPA == mid) |>
        select(-RPA) |> right_join(template, by = c('AgeCat', 'HHtype')) |> setDT()
      target[AgeCat %in% c('x == 1', 'x == 2', 'x == 3'), `:=`(BASELINE =
                                                                 0,
                                                               INTER = 0,
                                                               TARGET = 0)]
      target[is.na(BASELINE), `:=`(BASELINE = 0,
                                   INTER = 0,
                                   TARGET = 0)]
      setorder(target, AgeCat, HHtype)
      fwrite(target, paste0(inpath, '/AgeHHder_HHtype.csv'))
      
      template2 <- fread(paste0(workpath, 'Templates/age_tenure_template.csv'))
      target2 <- fread(
        paste0(
          ctrltotal.path,
          "tenure.targetfile.",
          reb,
          ".",
          scen,
          '.hdshp.',
          hdshp,
          ".y",
          yr,
          ".csv"
        )
      ) |>
        filter(RPA == mid) |>
        select(-RPA) |>
        right_join(template2, by = c('AgeCat', 'tenure')) |>
        setDT()
      
      target2[is.na(BASELINE), `:=`(BASELINE = 0,
                                    INTER = 0,
                                    TARGET = 0)]
      setorder(target2, AgeCat, tenure)
      fwrite(target2, paste0(inpath, '/age_tenure.csv'))
      
      message(paste0(mid, ' targets set'))
      
      pums <- fread(paste0(inpath, 'PUMS2021_', scen, '_', mid, '_', yr, '.csv'))
      
      source(here('scripts/reweighter/Manage_Scenario_RPA.R'))
      message(paste0(mid, ' reweighter complete'))
      
      mess <- data.table(txt = paste0(mid, ' reweighter complete'))
      fwrite(mess, paste0(outpath, mid, '_', yr, '_complete.txt'))
    }
  }
}

```

# 5 - Create reweighted PUMS files and QC the reweighter Outputs

# 5.0 - Check that all years/RPAs successfully converged

```{r}
  
log.comp <- c()
for (mid in mpos) {
  outpath <- paste0(workpath, scen, '/Output_Files/', mid, '/')
  logs <- list.files(path = outpath, pattern ='logfile*')
  for (f in logs){
    df <- readLines(paste0(outpath,f))
    log.comp <- c(log.comp, if (length(
      grep("Maximum deviation from target stopping condition achieved", df)
    ) != 0) {
      print(paste0(f, ' success'))
    } else if (length(grep("Maximum number of iterations reached", df)) !=
               0) {
      print(paste0(f, ' failed to converge'))
    } else if (length(grep(
      "No change in weights stopping condition met", df
    )) != 0) {
      print(paste0(f, ' no change in weights'))
    })
  }
}
fwrite(as.data.table(log.comp),paste0(workpath,scen,'/reweighter_convergence_check.txt'))




```

# 5.1 - Scenario 1

```{r}

comp.dt <- data.table()
if(scen=='Scenario_1'){
for (yr in 2020){
  for (mid in mpos) {
    inpath <- paste0(workpath, scen, '/Input_Files/', mid, '/')
    outpath <- paste0(workpath, scen, '/Output_Files/', mid, '/')
    # Summarize total households by mpo from target inputs to reweighter
    target <- fread(paste0(ctrltotal.path, 'type.targetfile.', reb, '.', scen, '.hdshp.', hdshp, '.y', yr, '.csv')) |>
      filter(RPA == mid) |> setDT()
    
    tsum <- target[, lapply(.SD, sum), RPA, .SDcols = 'TARGET']
    tsum[, TARGET := round(TARGET)][,year:=yr]
    setnames(tsum, 'RPA', 'mpo')
    setkey(tsum, mpo)
    
    # Calculate total households from the reweighter outputs
    pums <- fread(paste0(inpath, 'PUMS', vintage, '_', mid, '.csv'))
    setkey(pums, SERIALNO, SPORDER)

    x <- fread(paste0(outpath, 'weights_final_', scen, '_', mid, '_', yr, '.csv')) |> select(SERIALNO, SPORDER, new_WGTP) |> mutate(mpo = mid) |> setDT()
    setkey(x, SERIALNO, SPORDER)
    xp <- pums[x]
    fwrite(xp, paste0(outpath, '/reweighted_PUMS_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr, '.csv'))
    rw <- xp[SPORDER == 1, lapply(.SD, sum), .(mpo), .SDcols = c('WGTP', 'new_WGTP')]
    rw[, rw_HH := round(new_WGTP)][, pums_baseline_hh := round(WGTP)]
    setkey(rw, mpo)
    comp <- tsum[rw]
    comp.out <- comp[,.(year,mpo,pums_baseline_hh, TARGET,rw_HH)]
    comp.out[,diff:= TARGET - rw_HH]
    fwrite(comp.out,paste0(outpath,'reweighter_QC_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr,' .csv'))
    comp.dt <- rbind(comp.dt,comp.out)
  }
}
for (yr in c(2025,2030,2035,2040,2045,2050)){
  for (mid in mpos) {
    inpath <- paste0(workpath, scen, '/Input_Files/', mid, '/')
    outpath <- paste0(workpath, scen, '/Output_Files/', mid, '/')
    # Summarize total households by mpo from target inputs to reweighter
    target <- fread(paste0(ctrltotal.path, 'hhtype.targetfile.', reb, '.', scen, '.hdshp.', hdshp, '.y', yr, '.csv')) |>
      filter(RPA == mid) |> setDT()
    
    tsum <- target[, lapply(.SD, sum), RPA, .SDcols = 'TARGET']
    tsum[, TARGET := round(TARGET)][,year:=yr]
    setnames(tsum, 'RPA', 'mpo')
    setkey(tsum, mpo)
    
    # Calculate total households from the reweighter outputs
    pums <- fread(paste0(inpath, 'PUMS', vintage, '_', mid, '.csv'))
    setkey(pums, SERIALNO, SPORDER)

    x <- fread(paste0(outpath, 'weights_final_', scen, '_', mid, '_', yr, '.csv')) |> select(SERIALNO, SPORDER, new_WGTP) |> mutate(mpo = mid) |> setDT()
    setkey(x, SERIALNO, SPORDER)
    xp <- pums[x]
    fwrite(xp, paste0(outpath, '/reweighted_PUMS_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr, '.csv'))
    rw <- xp[SPORDER == 1, lapply(.SD, sum), .(mpo), .SDcols = c('WGTP', 'new_WGTP')]
    rw[, rw_HH := round(new_WGTP)][, pums_baseline_hh := round(WGTP)]
    setkey(rw, mpo)
    comp <- tsum[rw]
    comp.out <- comp[,.(year,mpo,pums_baseline_hh, TARGET,rw_HH)]
    comp.out[,diff:= TARGET - rw_HH]
    fwrite(comp.out,paste0(outpath,'reweighter_QC_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr,' .csv'))
    comp.dt <- rbind(comp.dt,comp.out)

  }
}

fwrite(comp.dt,paste0(workpath, scen, '/Output_Files/reweighter_QC_all_RPAs_all_years_', scen, '_', hdshp,' .csv'))

comp.dt[abs(diff)>1]
}
```

# 5.2 - Scenario 2/3/4

```{r}
c0 <- data.table()
comp.dt <- data.table()
if(scen %in% c('Scenario_2','Scenario_3','Scenario_4')){
for (yr in c(2020)){
  for (mid in mpos) {
    inpath <- paste0(workpath, scen, '/Input_Files/', mid, '/')
    outpath <- paste0(workpath, scen, '/Output_Files/', mid, '/')
    
    # Use Scenario 1 outputs for 2020 
    file.copy(paste0(workpath, 'Scenario_1/Output_Files/', mid, '/reweighted_PUMS_Scenario_1_', reb, '_', mid, '_acs1721_', yr, '.csv'),paste0(outpath, '/reweighted_PUMS_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr, '.csv'))
  }
}

for (yr in c(2025,2030,2035,2040,2045,2050)){
  for (mid in mpos) {
    inpath <- paste0(workpath, scen, '/Input_Files/', mid, '/')
    outpath <- paste0(workpath, scen, '/Output_Files/', mid, '/')
    
    
    # Calculate total households from the reweighter outputs
  
      pums <- fread(paste0(inpath, 'PUMS', vintage, '_', scen, '_', mid, '_', yr, '.csv'))
    setkey(pums, SERIALNO, SPORDER)

    x <- fread(paste0(outpath, 'weights_final_', scen, '_', mid, '_', yr, '.csv')) |> select(SERIALNO, SPORDER, new_WGTP) |> mutate(mpo = mid) |> setDT()
    setkey(x, SERIALNO, SPORDER)
    xp <- pums[x]
    fwrite(xp, paste0(outpath, 'reweighted_PUMS_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr, '.csv'))
    rw <- xp[SPORDER == 1, lapply(.SD, sum), .(mpo), .SDcols = c('WGTP', 'new_WGTP')]
    rw[, rw_HH := round(new_WGTP)][, pums_baseline_hh := round(WGTP)]
    setkey(rw, mpo)
    
    r1 <- xp[SPORDER == 1, lapply(.SD, sum), .(mpo,AgeCat,HHtype), .SDcols = c('WGTP', 'new_WGTP')]
    r1[, new_WGTP := round(new_WGTP)][, WGTP := round(WGTP)]
    r1[,AgeCat:=paste0('x == ',AgeCat)]
    r1[,HHtype:=paste0('x == ',HHtype)]
    setkey(r1, mpo,AgeCat,HHtype)
    # Summarize total households by mpo from target inputs to reweighter
    target <- fread(paste0(ctrltotal.path, 'hhtype.targetfile.', reb, '.', scen, '.hdshp.', hdshp, '.y', yr, '.csv')) |>
      filter(RPA == mid) |> rename(mpo=RPA) |> setDT()
    setkey(target,mpo,AgeCat,HHtype)
    
    cc <- r1[target]
    cc[,diff:=new_WGTP-TARGET]
    
    
    tsum <- target[, lapply(.SD, sum), mpo, .SDcols = 'TARGET']
    tsum[, TARGET := round(TARGET)][,year:=yr]
    #setnames(tsum, 'RPA', 'mpo')
    setkey(tsum, mpo)
    
    comp <- tsum[rw]
    comp.out <- comp[,.(year,mpo,pums_baseline_hh, TARGET,rw_HH)]
    comp.out[,diff:= TARGET - rw_HH]
    fwrite(comp.out,paste0(outpath,'reweighter_QC_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr,'.csv'))
    comp.dt <- rbind(comp.dt,comp.out)
    c0 <- rbind(c0,cc)
  }
}

fwrite(comp.dt,paste0(workpath, scen, '/Output_Files/reweighter_QC_all_RPAs_all_years_', scen, '_', hdshp,'.csv'))
fwrite(c0,paste0(workpath, scen, '/Output_Files/reweighter_QC_all_RPAs_by_AgeCat_HHtype_all_years_', scen, '_', hdshp,'.csv'))

comp.dt[abs(diff)>1]
c0[diff!=0]
}
```

# 6 - Create summary tables of reweighted households

```{r}

for (yr in c(2020,2025,2030,2035,2040,2045,2050)){
  sw <- data.table()
  for (mid in mpos){
    inp <- paste0(workpath, scen, '/Output_Files/', mid, '/')
    dt <- fread(paste0(inp, 'reweighted_PUMS_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr, '.csv'))
    dt[, year:=yr][,reb:=reb][,hdshp:=hdshp][,scen:=scen]
    sw <- rbind(sw,dt)
  }
 fwrite(sw,paste0(workpath,scen,'/Output_Files/reweighted_PUMS_', scen, '_', reb, '_', hdshp, '_', yr, '.csv'))
 }
```

# 7 - Vacancy and Unit Demand post-model calculations

# 7.1 - Scenario 1

```{r}

if(scen=='Scenario_1'){
# Set residential vacancy rate targets
# Load in the actual 2020 unit counts and vacancy rates from Census 2020 DHC

munis <- mapcdatakeys::all_muni_data_keys |>
  select(muni_id, muni_name, cosub_cn20, cmtyp08, mapc, mpo, rpa_acr) |>
  mutate(GEOID = as.character(cosub_cn20)) 

vars <- load_variables(2020, 'dhc') |> setDT()
vs <- vars[grepl('VACANCY', concept), name]
ns <- vars[grepl('VACANCY', concept), gsub(':', '', gsub(' ', '_', (gsub(
  ',', '', gsub(':!!', '_', (gsub(' !!', '', (
    label
  ))))
))))]

ten <- c('H10_002N','H10_010N') # Vacant owner-occupied units, Vacant renter-occupied units

vs <- c('H1_001N', 'P20_001N', vs, ten)

d1 <-
  get_decennial(
    year = 2020,
    sumfile = 'dhc',
    state = 'MA',
    geography = 'county subdivision',
    variables = vs
  ) |>
  pivot_wider(names_from = 'variable') |>
  select(-NAME) |>
  setDT()
setnames(d1, c('GEOID', 'total_units', 'total_households', gsub('Total','Vacant',ns), 'own_occ_units', 'rent_occ_units'))

m2 <- munis |> select(GEOID, muni_id, muni_name, mpo, rpa_acr, cmtyp08)
dhc <- d1 |>
  left_join(m2, by = 'GEOID') |>
  filter(!is.na(muni_id)) |> 
  mutate(
    total_available_units = total_households + Vacant_For_rent + Vacant_For_sale_only,
    total_unavailable_units = total_units - total_available_units,
    total_available_rental_units = Vacant_For_rent + rent_occ_units,
    total_available_owner_units = Vacant_For_sale_only + own_occ_units,
    rent_ratio = total_available_rental_units / total_available_units,
    own_ratio = total_available_owner_units / total_available_units,
    total_unavailable_rental_units = round(total_unavailable_units * rent_ratio),
    total_unavailable_owner_units = round(total_unavailable_units * own_ratio),
    total_rental_units = total_available_rental_units + total_unavailable_rental_units,
    total_owner_units = total_available_owner_units + total_unavailable_owner_units,
    total_available_vacant_units = Vacant_For_sale_only + Vacant_For_rent,
    vac_rate_avail = round(total_available_vacant_units / total_available_units, 3),
    rent_vac_rate_avail = round(Vacant_For_rent / total_available_rental_units,3),
    own_vac_rate_avail = round(Vacant_For_sale_only / total_available_owner_units,3),
    vac_rate_total = round(total_available_vacant_units / total_units, 3),
    rent_vac_rate_total = round(Vacant_For_rent / total_rental_units,3),
    own_vac_rate_total = round(Vacant_For_sale_only / total_owner_units,3)
    ) |> 
  rename(total_vacant_units=Vacant,
         vacant_available_to_rent=Vacant_For_rent,
         vacant_rented_not_occ=Vacant_Rented_not_occupied,
         vacant_available_to_own=Vacant_For_sale_only,
         vacant_sold_not_occ=Vacant_Sold_not_occupied,
         vacant_seasonal=Vacant_For_seasonal_recreational_or_occasional_use,
         vacant_migrant=Vacant_For_migrant_workers,
         vacant_other=Vacant_Other_vacant)
 
cs <- c(
  'total_units',
  'total_households',
  'total_vacant_units',
  'total_available_units',
  'total_available_vacant_units',
  'own_occ_units',
  'rent_occ_units',
  'vacant_sold_not_occ',
  'vacant_rented_not_occ',
  'total_rental_units',
  'total_owner_units',
  'total_available_rental_units',
  'vacant_available_to_rent',
  'total_available_owner_units',
  'vacant_available_to_own'
)


dsum101 <- dhc[,lapply(.SD,sum),rpa_acr,.SDcols=(cs)] |> 
  mutate(
    vac_rate_avail = round(total_available_vacant_units / total_available_units, 3),
    rent_vac_rate_avail = round(vacant_available_to_rent / total_available_rental_units,3),
    own_vac_rate_avail = round(vacant_available_to_own / total_available_owner_units,3),
    vac_rate_total = round(total_available_vacant_units / total_units, 3),
    rent_vac_rate_total = round(vacant_available_to_rent / total_rental_units,3),
    own_vac_rate_total = round(vacant_available_to_own / total_owner_units,3),
    all_unit_rental_vac_rate = round(1 - (rent_occ_units / total_rental_units),3),
    all_unit_owner_vac_rate = round(1 - (own_occ_units / total_owner_units),3)
    )
fwrite(dsum101,paste0(workpath,'MAPC101_all_RPA_available_vacancy_rates_by_tenure.csv'))

vax <- dsum101[,.(rpa_acr,total_units, total_households, own_occ_units,rent_occ_units,total_owner_units, total_available_owner_units, vacant_available_to_own, total_rental_units, total_available_rental_units, vacant_available_to_rent, own_vac_rate_total, rent_vac_rate_total, vac_rate_total)]

# Load reweighted future households data
dt <- data.table()

for (yr in c(2020,2025,2030,2035,2040,2045,2050)){
  tmp <- fread(paste0(workpath,scen,'/Output_Files/reweighted_PUMS_', scen, '_', reb, '_',hdshp,'_', yr, '.csv')) |> 
    select(year,reb,hdshp,scen,mpo,SERIALNO,SPORDER,AgeCat,tenure,TEN,new_WGTP)
  dt <- rbind(dt,tmp)
}
  
  hh.ten <- dt |> 
    filter(SPORDER==1) |> 
    mutate(tenure = case_when(
      TEN %in% 1:2 ~ 'own',
      TEN %in% 3:4 ~ 'rent')) |> 
    group_by(year,mpo,AgeCat,tenure) |> 
    count(wt=new_WGTP, name='hh') |> 
    ungroup() |> 
    # mutate(hh = round(hh)) |> 
    pivot_wider(names_from = c(tenure,year), id_cols = c(mpo,AgeCat), values_from = hh) |> 
    mutate(AgeCat = paste0('x == ',AgeCat)) |>
    setDT()
  setkey(hh.ten,mpo,AgeCat)
  
  ten.r <- fread(paste0(ctrltotal.path, "tenure_rates.pums.",hdshp,".csv"))
  
  setnames(ten.r,'rpa_acr','mpo')
  
  ten.r[tenure=='x == 1',tenure:='own']
  ten.r[tenure=='x == 2',tenure:='rent']
  
  ten.rw <- ten.r |> pivot_wider(names_from = tenure, id_cols = c(mpo,AgeCat), values_from = freq) %>% setDT()
  ten.rw[is.na(ten.rw)] <- 0
  ten.rw[,own:=round(own,2)][,rent:=round(rent,2)]
  ten.rw[,age_range:=rep(c('15-19','20-24','25-29','30-34','35-39','40-44','45-49','50-54','55-59','60-64','65-69','70-74','75-79','80-84','85+'),13)]
  setnames(ten.rw,c('mpo','AgeCat','owner share acs2021','renter share acs2021','Age Range'))
  setkey(ten.rw,mpo,AgeCat)
  
hh.ten.age <- ten.rw[hh.ten]
hh.ten.age[,Scenario:=scen]
hh.ten.age <- hh.ten.age |> relocate(`Age Range`,.after=AgeCat) |> relocate(Scenario,.before=mpo)
hh.ten.age[is.na(hh.ten.age)] <- 0

  hh <- dt |> 
    filter(SPORDER==1) |> 
    mutate(tenure = case_when(
      TEN %in% 1:2 ~ 'own',
      TEN %in% 3:4 ~ 'rent')) |> 
    group_by(year,mpo,tenure) |> 
    count(wt=new_WGTP, name='hh') |> 
    ungroup() |> 
    # mutate(hh = round(hh)) |> 
    setDT()
  uw <- hh |> 
    pivot_wider(id_cols = 'mpo', names_from = c('tenure','year'), values_from = 'hh') |>
    left_join(vax, by=c('mpo'='rpa_acr')) |> 
    setDT()
nm <- names(uw)[grepl('_20',names(uw))]
setnames(uw,nm,paste0('households_',nm))

uw[,rent_target:=rent_vac_rate_total][,own_target:=own_vac_rate_total]

uw[, own_units_needed_in_2025 := round(households_own_2025 / (1 - own_target))]
uw[, own_units_needed_in_2035 := round(households_own_2035 / (1 - own_target))]
uw[, own_units_needed_in_2050 := round(households_own_2050 / (1 - own_target))]
uw[, rent_units_needed_in_2025 := round(households_rent_2025 / (1 - rent_target))]
uw[, rent_units_needed_in_2035 := round(households_rent_2035 / (1 - rent_target))]
uw[, rent_units_needed_in_2050 := round(households_rent_2050 / (1 - rent_target))]

uw[,new_owner_units_needed_2020_to_2025 := own_units_needed_in_2025 - total_available_owner_units]
uw[,new_rental_units_needed_2020_to_2025 := rent_units_needed_in_2025 - total_available_rental_units]

uw[,new_owner_units_needed_2025_to_2035 := own_units_needed_in_2035 - own_units_needed_in_2025]
uw[,new_rental_units_needed_2025_to_2035 := rent_units_needed_in_2035 - rent_units_needed_in_2025]

uw[,new_owner_units_needed_2035_to_2050 := own_units_needed_in_2050 - own_units_needed_in_2035]
uw[,new_rental_units_needed_2035_to_2050 := rent_units_needed_in_2050 - rent_units_needed_in_2035]

uw[,new_owner_units_needed_2020_to_2050 := own_units_needed_in_2050 - total_available_owner_units]
uw[,new_rental_units_needed_2020_to_2050 := rent_units_needed_in_2050 - total_available_rental_units]

uw[,total_new_units_needed_2020_to_2025 := new_owner_units_needed_2020_to_2025+new_rental_units_needed_2020_to_2025]

uw[,total_new_units_needed_2025_to_2035 := new_owner_units_needed_2025_to_2035+new_rental_units_needed_2025_to_2035]

uw[,total_new_units_needed_2035_to_2050 := new_owner_units_needed_2035_to_2050+new_rental_units_needed_2035_to_2050]
uw[,scenario:=scen]

uw[,total_new_units_needed_2020_to_2050 := new_owner_units_needed_2020_to_2050+new_rental_units_needed_2020_to_2050]
uw[,scenario:=scen]

setnames(uw,'vac_rate_total','combined_vacancy_rate_2020')

outfile <- uw[,.(scenario,mpo,total_units,total_owner_units,total_rental_units,total_available_owner_units,total_available_rental_units,vacant_available_to_own,vacant_available_to_rent,own_vac_rate_total,rent_vac_rate_total,own_target,rent_target,combined_vacancy_rate_2020,own_occ_units,rent_occ_units,households_own_2020,households_rent_2020,households_own_2025,households_rent_2025,households_own_2030,households_rent_2030,households_own_2035,households_rent_2035,households_own_2040,households_rent_2040,households_own_2045,households_rent_2045,households_own_2050,households_rent_2050,own_units_needed_in_2025,own_units_needed_in_2035,own_units_needed_in_2050,rent_units_needed_in_2025,rent_units_needed_in_2035,rent_units_needed_in_2050,new_owner_units_needed_2020_to_2025,new_rental_units_needed_2020_to_2025,new_owner_units_needed_2025_to_2035,new_rental_units_needed_2025_to_2035
,new_owner_units_needed_2035_to_2050,new_rental_units_needed_2035_to_2050
,new_owner_units_needed_2020_to_2050,new_rental_units_needed_2020_to_2050,total_new_units_needed_2020_to_2025
,total_new_units_needed_2025_to_2035
,total_new_units_needed_2035_to_2050
,total_new_units_needed_2020_to_2050)]

setnames(outfile,c('own_target','rent_target','total_units','total_owner_units','total_rental_units','total_available_owner_units','total_available_rental_units','vacant_available_to_own','vacant_available_to_rent','own_vac_rate_total','rent_vac_rate_total','own_occ_units','rent_occ_units'),c('own_target_vacancy_rate','rent_target_vacancy_rate','total_units_2020','total_owner_units_2020','total_rental_units_2020','total_available_owner_units_2020','total_available_rental_units_2020','vacant_available_to_own_2020','vacant_available_to_rent_2020','owner_unit_vacancy_rate_2020','rental_unit_vacancy_rate_2020','Census 2020 owner-occupied households','Census 2020 renter-occupied households'))

setnames(outfile,names(outfile),gsub('_',' ',names(outfile)))
s2 <- copy(outfile)
s2[,mpo:=NULL]
swide <- s2[,lapply(.SD,sum,na.rm=T),scenario]
swide[,mpo:='Statewide']
of <- rbind(outfile,swide)
of[mpo=='Statewide',`owner unit vacancy rate 2020`:=`vacant available to own 2020` / `total available owner units 2020`]
of[mpo=='Statewide',`rental unit vacancy rate 2020`:=`vacant available to rent 2020` / `total available rental units 2020`]
of[mpo=='Statewide',`combined vacancy rate 2020`:=(`vacant available to own 2020` + `vacant available to rent 2020`) / (`total available owner units 2020` + `total available rental units 2020`)]
of[,`own target vacancy rate`:=target_own_rate]
of[,`rent target vacancy rate`:=target_rent_rate]
of[,`Change in Households 2020 - 2035`:=(`households own 2035` + `households rent 2035`) - (`households own 2020` + `households rent 2020`)]
of[,`Change in Households 2020 - 2050`:=(`households own 2050` + `households rent 2050`) - (`households own 2020` + `households rent 2020`)]


fwrite(of, paste0(workpath, scen, '/Output_Files/output.tbl_',scen,'_unit_demand_estimates.csv'))
fwrite(hh.ten.age,paste0(workpath, scen, '/Output_Files/output.tbl_',scen,'_households_by_age_by_tenure.csv'))

}
```

# 7.2 - Scenario 2 and 3

```{r}

if(scen %in% c('Scenario_2','Scenario_3')){

  shel <- fread(paste0(workpath,'shelter.hh_rpa.csv'))
  setnames(shel,c('mpo','hh.shel'))
  shel[,tenure:='rent']
  
  
  shel.age <- fread(paste0(workpath,'shelter.hh_rpa.agegroup.csv'))
  setnames(shel.age,c('mpo','age_group','hh.shel'))
  setkey(shel.age,mpo,age_group)
  
  
  target_rent_rate <- 0.074
  target_own_rate <- 0.015
  # Set residential vacancy rate targets
# Load in the actual 2020 unit counts and vacancy rates from Census 2020 DHC

munis <- mapcdatakeys::all_muni_data_keys |>
  select(muni_id, muni_name, cosub_cn20, cmtyp08, mapc, mpo, rpa_acr) |>
  mutate(GEOID = as.character(cosub_cn20)) 

vars <- load_variables(2020, 'dhc') |> setDT()
vs <- vars[grepl('VACANCY', concept), name]
ns <- vars[grepl('VACANCY', concept), gsub(':', '', gsub(' ', '_', (gsub(
  ',', '', gsub(':!!', '_', (gsub(' !!', '', (
    label
  ))))
))))]

ten <- c('H10_002N','H10_010N') # Vacant owner-occupied units, Vacant renter-occupied units

vs <- c('H1_001N', 'P20_001N', vs, ten)

d1 <-
  get_decennial(
    year = 2020,
    sumfile = 'dhc',
    state = 'MA',
    geography = 'county subdivision',
    variables = vs
  ) |>
  pivot_wider(names_from = 'variable') |>
  select(-NAME) |>
  setDT()
setnames(d1, c('GEOID', 'total_units', 'total_households', gsub('Total','Vacant',ns), 'own_occ_units', 'rent_occ_units'))

m2 <- munis |> select(GEOID, muni_id, muni_name, mpo, rpa_acr, cmtyp08)
dhc <- d1 |>
  left_join(m2, by = 'GEOID') |>
  filter(!is.na(muni_id)) |> 
  mutate(
    total_available_units = total_households + Vacant_For_rent + Vacant_For_sale_only,
    total_unavailable_units = total_units - total_available_units,
    total_available_rental_units = Vacant_For_rent + rent_occ_units,
    total_available_owner_units = Vacant_For_sale_only + own_occ_units,
    rent_ratio = total_available_rental_units / total_available_units,
    own_ratio = total_available_owner_units / total_available_units,
    total_unavailable_rental_units = round(total_unavailable_units * rent_ratio),
    total_unavailable_owner_units = round(total_unavailable_units * own_ratio),
    total_rental_units = total_available_rental_units + total_unavailable_rental_units,
    total_owner_units = total_available_owner_units + total_unavailable_owner_units,
    total_available_vacant_units = Vacant_For_sale_only + Vacant_For_rent,
    vac_rate_avail = round(total_available_vacant_units / total_available_units, 3),
    rent_vac_rate_avail = round(Vacant_For_rent / total_available_rental_units,3),
    own_vac_rate_avail = round(Vacant_For_sale_only / total_available_owner_units,3),
    vac_rate_total = round(total_available_vacant_units / total_units, 3),
    rent_vac_rate_total = round(Vacant_For_rent / total_rental_units,3),
    own_vac_rate_total = round(Vacant_For_sale_only / total_owner_units,3)
    ) |> 
  rename(total_vacant_units=Vacant,
         vacant_available_to_rent=Vacant_For_rent,
         vacant_rented_not_occ=Vacant_Rented_not_occupied,
         vacant_available_to_own=Vacant_For_sale_only,
         vacant_sold_not_occ=Vacant_Sold_not_occupied,
         vacant_seasonal=Vacant_For_seasonal_recreational_or_occasional_use,
         vacant_migrant=Vacant_For_migrant_workers,
         vacant_other=Vacant_Other_vacant)
 
cs <- c(
  'total_units',
  'total_households',
  'total_vacant_units',
  'total_available_units',
  'total_available_vacant_units',
  'own_occ_units',
  'rent_occ_units',
  'vacant_sold_not_occ',
  'vacant_rented_not_occ',
  'total_rental_units',
  'total_owner_units',
  'total_available_rental_units',
  'vacant_available_to_rent',
  'total_available_owner_units',
  'vacant_available_to_own'
)


dsum101 <- dhc[,lapply(.SD,sum),rpa_acr,.SDcols=(cs)] |> 
  mutate(
    vac_rate_avail = round(total_available_vacant_units / total_available_units, 3),
    rent_vac_rate_avail = round(vacant_available_to_rent / total_available_rental_units,3),
    own_vac_rate_avail = round(vacant_available_to_own / total_available_owner_units,3),
    vac_rate_total = round(total_available_vacant_units / total_units, 3),
    rent_vac_rate_total = round(vacant_available_to_rent / total_rental_units,3),
    own_vac_rate_total = round(vacant_available_to_own / total_owner_units,3),
    all_unit_rental_vac_rate = round(1 - (rent_occ_units / total_rental_units),3),
    all_unit_owner_vac_rate = round(1 - (own_occ_units / total_owner_units),3)
    )
fwrite(dsum101,paste0(workpath,'MAPC101_all_RPA_available_vacancy_rates_by_tenure.csv'))

vax <- dsum101[,.(rpa_acr,total_units, total_households, own_occ_units,rent_occ_units,total_owner_units, total_available_owner_units, vacant_available_to_own, total_rental_units, total_available_rental_units, vacant_available_to_rent, own_vac_rate_total, rent_vac_rate_total, vac_rate_total)]

# Load reweighted future households data
dt <- data.table()

for (yr in c(2020,2025,2030,2035,2040,2045,2050)){
  tmp <- fread(paste0(workpath,scen,'/Output_Files/reweighted_PUMS_', scen, '_', reb, '_',hdshp,'_', yr, '.csv')) |> 
    select(year,reb,hdshp,scen,mpo,SERIALNO,SPORDER,AgeCat,tenure,TEN,new_WGTP)
  dt <- rbind(dt,tmp)
}
  
  hh.ten <- dt |> 
    filter(SPORDER==1) |> 
    mutate(tenure = case_when(
      TEN %in% 1:2 ~ 'own',
      TEN %in% 3:4 ~ 'rent')) |> 
    group_by(year,mpo,AgeCat,tenure) |> 
    count(wt=new_WGTP, name='hh') |> 
    ungroup() |> 
    # mutate(hh = round(hh)) |> 
    pivot_wider(names_from = c(tenure,year), id_cols = c(mpo,AgeCat), values_from = hh) |> 
    mutate(AgeCat = paste0('x == ',AgeCat)) |>
    setDT()
  setkey(hh.ten,mpo,AgeCat)
  
  ten.r <- fread(paste0(ctrltotal.path, "tenure_rates.pums.",hdshp,".csv"))
  setnames(ten.r,'rpa_acr','mpo')
  ten.r[tenure=='x == 1',tenure:='own']
  ten.r[tenure=='x == 2',tenure:='rent']
  
  ten.rw <- ten.r |> pivot_wider(names_from = tenure, id_cols = c(mpo,AgeCat), values_from = freq) %>% setDT()
  ten.rw[is.na(ten.rw)] <- 0
  ten.rw[,own:=round(own,2)][,rent:=round(rent,2)]
  ten.rw[,age_group:=rep(c('15-19','20-24','25-29','30-34','35-39','40-44','45-49','50-54','55-59','60-64','65-69','70-74','75-79','80-84','85+'),13)]
  
  setkey(ten.rw,mpo,age_group)
  ten.rw <- shel.age[ten.rw]
  ten.rw[is.na(ten.rw)] <- 0
  
  setnames(ten.rw,c('mpo','Age Range','hh.shel','AgeCat','owner share acs2021','renter share acs2021'))
  setkey(ten.rw,mpo,AgeCat)
  setkey(hh.ten,mpo,AgeCat)
  
hh.ten.age <- ten.rw[hh.ten]
hh.ten.age <- hh.ten.age |> 
  mutate(Scenario = scen,
         rent_2030 = rent_2030 + hh.shel,
         rent_2035 = rent_2035 + hh.shel,
         rent_2040 = rent_2040 + hh.shel,
         rent_2045 = rent_2045 + hh.shel,
         rent_2050 = rent_2050 + hh.shel
         ) |> 
  relocate(`Age Range`,.after=AgeCat) |> 
  relocate(own_2020,.before=rent_2020) |> 
  relocate(own_2025,.before=rent_2025) |> 
  relocate(Scenario,.before=mpo)

hh.ten.age[is.na(hh.ten.age)] <- 0


  hh <- dt |> 
    filter(SPORDER==1) |> 
    mutate(tenure = case_when(
      TEN %in% 1:2 ~ 'own',
      TEN %in% 3:4 ~ 'rent')) |> 
    group_by(year,mpo,tenure) |> 
    count(wt=new_WGTP, name='hh') |> 
    ungroup() |> 
    # mutate(hh = round(hh)) |> 
    setDT() 
  setkey(hh,mpo,tenure)  
  setkey(shel,mpo,tenure)
  
  hh <- shel[hh]
  hh[is.na(hh.shel),hh.shel:=0]
  hh[year>2025 & tenure=='rent',hh:=hh+hh.shel]
  hh[,hh.shel:=NULL]
  
  uw <- hh |> 
    pivot_wider(id_cols = 'mpo', names_from = c('tenure','year'), values_from = 'hh') |>
    left_join(vax, by=c('mpo'='rpa_acr')) |> 
    setDT()
nm <- names(uw)[grepl('_20',names(uw))]
setnames(uw,nm,paste0('households_',nm))

uw[,rent_target25:=rent_vac_rate_total][,own_target25:=own_vac_rate_total]
uw[,rent_target:=target_rent_rate][,own_target:=target_own_rate]

uw[, own_units_needed_in_2025 := round(households_own_2025 / (1 - own_target25))]
uw[, own_units_needed_in_2035 := round(households_own_2035 / (1 - own_target))]
uw[, own_units_needed_in_2050 := round(households_own_2050 / (1 - own_target))]
uw[, rent_units_needed_in_2025 := round(households_rent_2025 / (1 - rent_target25))]
uw[, rent_units_needed_in_2035 := round(households_rent_2035 / (1 - rent_target))]
uw[, rent_units_needed_in_2050 := round(households_rent_2050 / (1 - rent_target))]

uw[,new_owner_units_needed_2020_to_2025 := own_units_needed_in_2025 - total_available_owner_units]
uw[,new_rental_units_needed_2020_to_2025 := rent_units_needed_in_2025 - total_available_rental_units]

uw[,new_owner_units_needed_2025_to_2035 := own_units_needed_in_2035 - own_units_needed_in_2025]
uw[,new_rental_units_needed_2025_to_2035 := rent_units_needed_in_2035 - rent_units_needed_in_2025]

uw[,new_owner_units_needed_2035_to_2050 := own_units_needed_in_2050 - own_units_needed_in_2035]
uw[,new_rental_units_needed_2035_to_2050 := rent_units_needed_in_2050 - rent_units_needed_in_2035]

uw[,new_owner_units_needed_2020_to_2050 := own_units_needed_in_2050 - total_available_owner_units]
uw[,new_rental_units_needed_2020_to_2050 := rent_units_needed_in_2050 - total_available_rental_units]

uw[,total_new_units_needed_2020_to_2025 := new_owner_units_needed_2020_to_2025+new_rental_units_needed_2020_to_2025]

uw[,total_new_units_needed_2025_to_2035 := new_owner_units_needed_2025_to_2035+new_rental_units_needed_2025_to_2035]

uw[,total_new_units_needed_2035_to_2050 := new_owner_units_needed_2035_to_2050+new_rental_units_needed_2035_to_2050]
uw[,scenario:=scen]

uw[,total_new_units_needed_2020_to_2050 := new_owner_units_needed_2020_to_2050+new_rental_units_needed_2020_to_2050]
uw[,scenario:=scen]

setnames(uw,'vac_rate_total','combined_vacancy_rate_2020')

outfile <- uw[,.(scenario,mpo,total_units,total_owner_units,total_rental_units,total_available_owner_units,total_available_rental_units,vacant_available_to_own,vacant_available_to_rent,own_vac_rate_total,rent_vac_rate_total,own_target,rent_target,combined_vacancy_rate_2020,own_occ_units,rent_occ_units,households_own_2020,households_rent_2020,households_own_2025,households_rent_2025,households_own_2030,households_rent_2030,households_own_2035,households_rent_2035,households_own_2040,households_rent_2040,households_own_2045,households_rent_2045,households_own_2050,households_rent_2050,own_units_needed_in_2025,own_units_needed_in_2035,own_units_needed_in_2050,rent_units_needed_in_2025,rent_units_needed_in_2035,rent_units_needed_in_2050,new_owner_units_needed_2020_to_2025,new_rental_units_needed_2020_to_2025,new_owner_units_needed_2025_to_2035,new_rental_units_needed_2025_to_2035
,new_owner_units_needed_2035_to_2050,new_rental_units_needed_2035_to_2050
,new_owner_units_needed_2020_to_2050,new_rental_units_needed_2020_to_2050,total_new_units_needed_2020_to_2025
,total_new_units_needed_2025_to_2035
,total_new_units_needed_2035_to_2050
,total_new_units_needed_2020_to_2050)]

setnames(outfile,c('own_target','rent_target','total_units','total_owner_units','total_rental_units','total_available_owner_units','total_available_rental_units','vacant_available_to_own','vacant_available_to_rent','own_vac_rate_total','rent_vac_rate_total','own_occ_units','rent_occ_units'),c('own_target_vacancy_rate','rent_target_vacancy_rate','total_units_2020','total_owner_units_2020','total_rental_units_2020','total_available_owner_units_2020','total_available_rental_units_2020','vacant_available_to_own_2020','vacant_available_to_rent_2020','owner_unit_vacancy_rate_2020','rental_unit_vacancy_rate_2020','Census 2020 owner-occupied households','Census 2020 renter-occupied households'))

setnames(outfile,names(outfile),gsub('_',' ',names(outfile)))
s2 <- copy(outfile)
s2[,mpo:=NULL]
swide <- s2[,lapply(.SD,sum,na.rm=T),scenario]
swide[,mpo:='Statewide']
of <- rbind(outfile,swide)
of[mpo=='Statewide',`owner unit vacancy rate 2020`:=`vacant available to own 2020` / `total available owner units 2020`]
of[mpo=='Statewide',`rental unit vacancy rate 2020`:=`vacant available to rent 2020` / `total available rental units 2020`]
of[mpo=='Statewide',`combined vacancy rate 2020`:=(`vacant available to own 2020` + `vacant available to rent 2020`) / (`total available owner units 2020` + `total available rental units 2020`)]
of[,`own target vacancy rate`:=target_own_rate]
of[,`rent target vacancy rate`:=target_rent_rate]
of[,`Change in Households 2020 - 2035`:=(`households own 2035` + `households rent 2035`) - (`households own 2020` + `households rent 2020`)]
of[,`Change in Households 2020 - 2050`:=(`households own 2050` + `households rent 2050`) - (`households own 2020` + `households rent 2020`)]


fwrite(of, paste0(workpath, scen, '/Output_Files/output.tbl_',scen,'_unit_demand_estimates.csv'))
fwrite(hh.ten.age,paste0(workpath, scen, '/Output_Files/output.tbl_',scen,'_households_by_age_by_tenure.csv'))

}


```

# 7.2 - Scenario 4

```{r}

if(scen %in% c('Scenario_4')){

  target_rent_rate <- 0.074
  target_own_rate <- 0.015
  # Set residential vacancy rate targets
# Load in the actual 2020 unit counts and vacancy rates from Census 2020 DHC

munis <- mapcdatakeys::all_muni_data_keys |>
  select(muni_id, muni_name, cosub_cn20, cmtyp08, mapc, mpo, rpa_acr) |>
  mutate(GEOID = as.character(cosub_cn20)) 

vars <- load_variables(2020, 'dhc') |> setDT()
vs <- vars[grepl('VACANCY', concept), name]
ns <- vars[grepl('VACANCY', concept), gsub(':', '', gsub(' ', '_', (gsub(
  ',', '', gsub(':!!', '_', (gsub(' !!', '', (
    label
  ))))
))))]

ten <- c('H10_002N','H10_010N') # Vacant owner-occupied units, Vacant renter-occupied units

vs <- c('H1_001N', 'P20_001N', vs, ten)

d1 <-
  get_decennial(
    year = 2020,
    sumfile = 'dhc',
    state = 'MA',
    geography = 'county subdivision',
    variables = vs
  ) |>
  pivot_wider(names_from = 'variable') |>
  select(-NAME) |>
  setDT()
setnames(d1, c('GEOID', 'total_units', 'total_households', gsub('Total','Vacant',ns), 'own_occ_units', 'rent_occ_units'))

m2 <- munis |> select(GEOID, muni_id, muni_name, mpo, rpa_acr, cmtyp08)
dhc <- d1 |>
  left_join(m2, by = 'GEOID') |>
  filter(!is.na(muni_id)) |> 
  mutate(
    total_available_units = total_households + Vacant_For_rent + Vacant_For_sale_only,
    total_unavailable_units = total_units - total_available_units,
    total_available_rental_units = Vacant_For_rent + rent_occ_units,
    total_available_owner_units = Vacant_For_sale_only + own_occ_units,
    rent_ratio = total_available_rental_units / total_available_units,
    own_ratio = total_available_owner_units / total_available_units,
    total_unavailable_rental_units = round(total_unavailable_units * rent_ratio),
    total_unavailable_owner_units = round(total_unavailable_units * own_ratio),
    total_rental_units = total_available_rental_units + total_unavailable_rental_units,
    total_owner_units = total_available_owner_units + total_unavailable_owner_units,
    total_available_vacant_units = Vacant_For_sale_only + Vacant_For_rent,
    vac_rate_avail = round(total_available_vacant_units / total_available_units, 3),
    rent_vac_rate_avail = round(Vacant_For_rent / total_available_rental_units,3),
    own_vac_rate_avail = round(Vacant_For_sale_only / total_available_owner_units,3),
    vac_rate_total = round(total_available_vacant_units / total_units, 3),
    rent_vac_rate_total = round(Vacant_For_rent / total_rental_units,3),
    own_vac_rate_total = round(Vacant_For_sale_only / total_owner_units,3)
    ) |> 
  rename(total_vacant_units=Vacant,
         vacant_available_to_rent=Vacant_For_rent,
         vacant_rented_not_occ=Vacant_Rented_not_occupied,
         vacant_available_to_own=Vacant_For_sale_only,
         vacant_sold_not_occ=Vacant_Sold_not_occupied,
         vacant_seasonal=Vacant_For_seasonal_recreational_or_occasional_use,
         vacant_migrant=Vacant_For_migrant_workers,
         vacant_other=Vacant_Other_vacant)
 
cs <- c(
  'total_units',
  'total_households',
  'total_vacant_units',
  'total_available_units',
  'total_available_vacant_units',
  'own_occ_units',
  'rent_occ_units',
  'vacant_sold_not_occ',
  'vacant_rented_not_occ',
  'total_rental_units',
  'total_owner_units',
  'total_available_rental_units',
  'vacant_available_to_rent',
  'total_available_owner_units',
  'vacant_available_to_own'
)


dsum101 <- dhc[,lapply(.SD,sum),rpa_acr,.SDcols=(cs)] |> 
  mutate(
    vac_rate_avail = round(total_available_vacant_units / total_available_units, 3),
    rent_vac_rate_avail = round(vacant_available_to_rent / total_available_rental_units,3),
    own_vac_rate_avail = round(vacant_available_to_own / total_available_owner_units,3),
    vac_rate_total = round(total_available_vacant_units / total_units, 3),
    rent_vac_rate_total = round(vacant_available_to_rent / total_rental_units,3),
    own_vac_rate_total = round(vacant_available_to_own / total_owner_units,3),
    all_unit_rental_vac_rate = round(1 - (rent_occ_units / total_rental_units),3),
    all_unit_owner_vac_rate = round(1 - (own_occ_units / total_owner_units),3)
    )
fwrite(dsum101,paste0(workpath,'MAPC101_all_RPA_available_vacancy_rates_by_tenure.csv'))

vax <- dsum101[,.(rpa_acr,total_units, total_households, own_occ_units,rent_occ_units,total_owner_units, total_available_owner_units, vacant_available_to_own, total_rental_units, total_available_rental_units, vacant_available_to_rent, own_vac_rate_total, rent_vac_rate_total, vac_rate_total)]

# Load reweighted future households data
dt <- data.table()

for (yr in c(2020,2025,2030,2035,2040,2045,2050)){
  tmp <- fread(paste0(workpath,scen,'/Output_Files/reweighted_PUMS_', scen, '_', reb, '_',hdshp,'_', yr, '.csv')) |> 
    select(year,reb,hdshp,scen,mpo,SERIALNO,SPORDER,AgeCat,tenure,TEN,new_WGTP)
  dt <- rbind(dt,tmp)
}
  
  hh.ten <- dt |> 
    filter(SPORDER==1) |> 
    mutate(tenure = case_when(
      TEN %in% 1:2 ~ 'own',
      TEN %in% 3:4 ~ 'rent')) |> 
    group_by(year,mpo,AgeCat,tenure) |> 
    count(wt=new_WGTP, name='hh') |> 
    ungroup() |> 
    # mutate(hh = round(hh)) |> 
    pivot_wider(names_from = c(tenure,year), id_cols = c(mpo,AgeCat), values_from = hh) |> 
    mutate(AgeCat = paste0('x == ',AgeCat)) |>
    setDT()
  setkey(hh.ten,mpo,AgeCat)
  
  ten.r <- fread(paste0(ctrltotal.path, "tenure_rates.pums.",hdshp,".csv"))
  setnames(ten.r,'rpa_acr','mpo')
  ten.r[tenure=='x == 1',tenure:='own']
  ten.r[tenure=='x == 2',tenure:='rent']
  
  ten.rw <- ten.r |> pivot_wider(names_from = tenure, id_cols = c(mpo,AgeCat), values_from = freq) %>% setDT()
  ten.rw[is.na(ten.rw)] <- 0
  ten.rw[,own:=round(own,2)][,rent:=round(rent,2)]
  ten.rw[,age_group:=rep(c('15-19','20-24','25-29','30-34','35-39','40-44','45-49','50-54','55-59','60-64','65-69','70-74','75-79','80-84','85+'),13)]
  
  setkey(ten.rw,mpo,age_group)
  ten.rw[is.na(ten.rw)] <- 0
  ten.rw <- ten.rw |> relocate(age_group, .after=mpo)
  setnames(ten.rw,c('mpo','Age Range','AgeCat','owner share acs2021','renter share acs2021'))
  setkey(ten.rw,mpo,AgeCat)
  setkey(hh.ten,mpo,AgeCat)
  
hh.ten.age <- ten.rw[hh.ten]
hh.ten.age <- hh.ten.age |> 
  mutate(Scenario = scen) |> 
  relocate(`Age Range`,.after=AgeCat) |> 
  relocate(own_2020,.before=rent_2020) |> 
  relocate(own_2025,.before=rent_2025) |> 
  relocate(Scenario,.before=mpo)

hh.ten.age[is.na(hh.ten.age)] <- 0


  hh <- dt |> 
    filter(SPORDER==1) |> 
    mutate(tenure = case_when(
      TEN %in% 1:2 ~ 'own',
      TEN %in% 3:4 ~ 'rent')) |> 
    group_by(year,mpo,tenure) |> 
    count(wt=new_WGTP, name='hh') |> 
    ungroup() |> 
    # mutate(hh = round(hh)) |> 
    setDT() 
  setkey(hh,mpo,tenure)  
 
  uw <- hh |> 
    pivot_wider(id_cols = 'mpo', names_from = c('tenure','year'), values_from = 'hh') |>
    left_join(vax, by=c('mpo'='rpa_acr')) |> 
    setDT()
nm <- names(uw)[grepl('_20',names(uw))]
setnames(uw,nm,paste0('households_',nm))

uw[,rent_target25:=rent_vac_rate_total][,own_target25:=own_vac_rate_total]
uw[,rent_target:=target_rent_rate][,own_target:=target_own_rate]

uw[, own_units_needed_in_2025 := round(households_own_2025 / (1 - own_target25))]
uw[, own_units_needed_in_2035 := round(households_own_2035 / (1 - own_target))]
uw[, own_units_needed_in_2050 := round(households_own_2050 / (1 - own_target))]
uw[, rent_units_needed_in_2025 := round(households_rent_2025 / (1 - rent_target25))]
uw[, rent_units_needed_in_2035 := round(households_rent_2035 / (1 - rent_target))]
uw[, rent_units_needed_in_2050 := round(households_rent_2050 / (1 - rent_target))]

uw[,new_owner_units_needed_2020_to_2025 := own_units_needed_in_2025 - total_available_owner_units]
uw[,new_rental_units_needed_2020_to_2025 := rent_units_needed_in_2025 - total_available_rental_units]

uw[,new_owner_units_needed_2025_to_2035 := own_units_needed_in_2035 - own_units_needed_in_2025]
uw[,new_rental_units_needed_2025_to_2035 := rent_units_needed_in_2035 - rent_units_needed_in_2025]

uw[,new_owner_units_needed_2035_to_2050 := own_units_needed_in_2050 - own_units_needed_in_2035]
uw[,new_rental_units_needed_2035_to_2050 := rent_units_needed_in_2050 - rent_units_needed_in_2035]

uw[,new_owner_units_needed_2020_to_2050 := own_units_needed_in_2050 - total_available_owner_units]
uw[,new_rental_units_needed_2020_to_2050 := rent_units_needed_in_2050 - total_available_rental_units]

uw[,total_new_units_needed_2020_to_2025 := new_owner_units_needed_2020_to_2025+new_rental_units_needed_2020_to_2025]

uw[,total_new_units_needed_2025_to_2035 := new_owner_units_needed_2025_to_2035+new_rental_units_needed_2025_to_2035]

uw[,total_new_units_needed_2035_to_2050 := new_owner_units_needed_2035_to_2050+new_rental_units_needed_2035_to_2050]
uw[,scenario:=scen]

uw[,total_new_units_needed_2020_to_2050 := new_owner_units_needed_2020_to_2050+new_rental_units_needed_2020_to_2050]
uw[,scenario:=scen]

setnames(uw,'vac_rate_total','combined_vacancy_rate_2020')

outfile <- uw[,.(scenario,mpo,total_units,total_owner_units,total_rental_units,total_available_owner_units,total_available_rental_units,vacant_available_to_own,vacant_available_to_rent,own_vac_rate_total,rent_vac_rate_total,own_target,rent_target,combined_vacancy_rate_2020,own_occ_units,rent_occ_units,households_own_2020,households_rent_2020,households_own_2025,households_rent_2025,households_own_2030,households_rent_2030,households_own_2035,households_rent_2035,households_own_2040,households_rent_2040,households_own_2045,households_rent_2045,households_own_2050,households_rent_2050,own_units_needed_in_2025,own_units_needed_in_2035,own_units_needed_in_2050,rent_units_needed_in_2025,rent_units_needed_in_2035,rent_units_needed_in_2050,new_owner_units_needed_2020_to_2025,new_rental_units_needed_2020_to_2025,new_owner_units_needed_2025_to_2035,new_rental_units_needed_2025_to_2035
,new_owner_units_needed_2035_to_2050,new_rental_units_needed_2035_to_2050
,new_owner_units_needed_2020_to_2050,new_rental_units_needed_2020_to_2050,total_new_units_needed_2020_to_2025
,total_new_units_needed_2025_to_2035
,total_new_units_needed_2035_to_2050
,total_new_units_needed_2020_to_2050)]

setnames(outfile,c('own_target','rent_target','total_units','total_owner_units','total_rental_units','total_available_owner_units','total_available_rental_units','vacant_available_to_own','vacant_available_to_rent','own_vac_rate_total','rent_vac_rate_total','own_occ_units','rent_occ_units'),c('own_target_vacancy_rate','rent_target_vacancy_rate','total_units_2020','total_owner_units_2020','total_rental_units_2020','total_available_owner_units_2020','total_available_rental_units_2020','vacant_available_to_own_2020','vacant_available_to_rent_2020','owner_unit_vacancy_rate_2020','rental_unit_vacancy_rate_2020','Census 2020 owner-occupied households','Census 2020 renter-occupied households'))

setnames(outfile,names(outfile),gsub('_',' ',names(outfile)))
s2 <- copy(outfile)
s2[,mpo:=NULL]
swide <- s2[,lapply(.SD,sum,na.rm=T),scenario]
swide[,mpo:='Statewide']
of <- rbind(outfile,swide)
of[mpo=='Statewide',`owner unit vacancy rate 2020`:=`vacant available to own 2020` / `total available owner units 2020`]
of[mpo=='Statewide',`rental unit vacancy rate 2020`:=`vacant available to rent 2020` / `total available rental units 2020`]
of[mpo=='Statewide',`combined vacancy rate 2020`:=(`vacant available to own 2020` + `vacant available to rent 2020`) / (`total available owner units 2020` + `total available rental units 2020`)]
of[,`own target vacancy rate`:=target_own_rate]
of[,`rent target vacancy rate`:=target_rent_rate]
of[,`Change in Households 2020 - 2035`:=(`households own 2035` + `households rent 2035`) - (`households own 2020` + `households rent 2020`)]
of[,`Change in Households 2020 - 2050`:=(`households own 2050` + `households rent 2050`) - (`households own 2020` + `households rent 2020`)]

fwrite(of, paste0(workpath, scen, '/Output_Files/output.tbl_',scen,'_unit_demand_estimates.csv'))
fwrite(hh.ten.age,paste0(workpath, scen, '/Output_Files/output.tbl_',scen,'_households_by_age_by_tenure.csv'))

}


```

# 8 - Population Post-Processor

# 8.1 - Post-processor 02 - population difference calculation

```{r}
#| eval: false

# Load UMDI HHpop comparison data

umdi <- fread(paste0(ctrltotal.path,'pop.in.hhds.MAPC101_by.RPA.ageCAT6.csv'))

setnames(
  umdi,
  c(
    'mpo',
    'ageCAT6',
    'year',
    'pop',
    'umdi_hh_pop'
  )
)

umdi <-
  umdi[, lapply(.SD, sum, na.rm = T), .(year, mpo, ageCAT6), .SDcols = c('umdi_hh_pop')]
umdi[, umdi_hh_pop := round(umdi_hh_pop)]
umdi <- umdi[year > 2020]
setkey(umdi, mpo, year, ageCAT6)

##  CALCULATE differences in HHpop by age between UMDI/Control Totals and reweighter Outputs


for (yr in c(2035)){
    dx <- fread(paste0(workpath,scen,'/Output_Files/reweighted_PUMS_', scen, '_', reb, '_', hdshp, '_', yr, '.csv'))
    dx[, c("WGT1", "WGT2") := tstrsplit(new_WGTP, ".", fixed=TRUE)]
    dx[,WGT1:=as.numeric(WGT1)]
    dx[is.na(WGT2),WGT2:=0]
    dx[,WGT2:=as.numeric(paste0('0.',WGT2))]
    
    dfrac <- dx[WGT2>0]
    dfrac[,new_WGTP:=WGT2]
    dfrac[,hid:=paste0(SERIALNO,'_',mpo,'_frac')]
    
    xr <- dx[rep(1:.N,WGT1)]
    xr[,new_WGTP:=1]
    setorder(xr,SERIALNO,SPORDER)
    xr[,ss:=seq_along(.I),.(SERIALNO,SPORDER)]
    xr[, hid := paste(SERIALNO, mpo, ss, sep = '_')]
    xr[,ss:=NULL]
    dt <- rbind(xr,dfrac)
    dt[,pop:=new_WGTP]
    dt[,pid:=paste0(hid,'_',SPORDER)]
    dt[,person_num:=SPORDER][,age:=AGEP]
    fwrite(dt,paste0(output_path,'expanded_reweighted_PUMS_',yr,'.csv'))
    uspop <- dt[,lapply(.SD,sum),.(mpo,year,ageCAT6),.SDcols='pop']
    setkey(uspop, mpo, year, ageCAT6)
    adiff <- umdi[uspop]
    adiff[, diff := pop - umdi_hh_pop]
    setnames(adiff, 'pop', 'urbansim_hh_pop')
    fwrite(adiff, paste0(output_path,'umdi_pop_age_difference_',yr,'.csv'))
rm(xr,dt,dx,dfrac);gc()
      }



```

# 8.2 - Post-processor 03 - person shifts

```{r}
#| eval: false
  

for (yr in c(2035)){
  # adiff <- fread(paste0(output_path,'umdi_pop_age_difference_',yr,'.csv'))
  adiff <- fread(paste0(output_path,'umdi_pop_age_difference_',yr,'.csv'))
    
  adiff[,diff:=round(diff)]
  
  dt_shift <- data.table()
  #dt <- fread(paste0(output_path,'expanded_reweighted_PUMS_',yr,'.csv'))
  dt <- fread(paste0(output_path,'expanded_reweighted_PUMS_',yr,'.csv'))
  
  for (m in unique(adiff$mpo)) {
    
    swm <- dt[mpo==m]
    swm[,person_num:=SPORDER][,age:=AGEP]
    gc()
    diff <- adiff[mpo == m & year == yr]
    
    
    # ageCAT 2 and 3
    for (i in 2:3) {
      dlag <- diff[ageCAT6 == i - 1, diff]
      da <- diff[ageCAT6 == i, diff]
      dlead <- diff[ageCAT6 == i + 1, diff]
      
      if (da > 0 & dlag < 0 & dlead < 0) {
        if ((abs(dlag) + abs(dlead)) > da) {
          subN <- abs(round(da * (dlag / (dlag + dlead))))
          addN <- abs(round(da * (dlead / (dlag + dlead))))
        }
        else if ((abs(dlag) + abs(dlead)) <= da) {
          subN <- abs(dlag)
          addN <- abs(dlead)
        }
        tosub <-
          swm[ageCAT6 == i &
                person_num > 1, sample(unique(pid), subN)]
        toadd <-
          swm[ageCAT6 == i &
                person_num > 1, sample(unique(pid), addN)]
        swm[pid %in% tosub, `:=` (age = age - 5, ageCAT6 = ageCAT6 - 1)]
        swm[pid %in% toadd, `:=` (age = age + 5, ageCAT6 = ageCAT6 + 1)]
      }
      if (da > 0 & dlag < 0 & dlead > 0) {
        if (abs(dlag) > da) {
          subN <- abs(da)
        }
        else if (abs(dlag) <= da) {
          subN <- abs(dlag)
        }
        tosub <-
          swm[ageCAT6 == i &
                person_num > 1, sample(unique(pid), subN)]
        swm[pid %in% tosub, `:=` (age = age - 5, ageCAT6 = ageCAT6 - 1)]
      }
      if (da > 0 & dlag > 0 & dlead < 0) {
        if (abs(dlead) > da) {
          addN <- abs(da)
        }
        else if (abs(dlead) <= da) {
          addN <- abs(dlead)
        }
        toadd <-
          swm[ageCAT6 == i &
                person_num > 1, sample(unique(pid), addN)]
        swm[pid %in% toadd, `:=` (age = age + 5, ageCAT6 = ageCAT6 + 1)]
      }
      # Update age difference table
      newlag <- swm[ageCAT6 == i - 1, .N]
      newpop <- swm[ageCAT6 == i, .N]
      newlead <- swm[ageCAT6 == i + 1, .N]
      diff[ageCAT6 == i - 1, `:=`(urbansim_hh_pop = newlag, diff = newlag - umdi_hh_pop)]
      diff[ageCAT6 == i, `:=`(urbansim_hh_pop = newpop, diff = newpop - umdi_hh_pop)]
      diff[ageCAT6 == i + 1, `:=`(urbansim_hh_pop = newlead, diff = newlead - umdi_hh_pop)]
      suppressWarnings(rm(newlag,newpop,newlead,addN,subN,da,dlag,dlead,tosub,toadd))
      gc()
    }
    
    # ageCAT = 1
    da <- diff[ageCAT6 == 1, diff]
    dlead <- diff[ageCAT6 == 2, diff]
    if (da > 0 & dlead < 0) {
      if (abs(dlead) > da) {
        addN <- abs(da)
      }
      else if (abs(dlead) <= da) {
        addN <- abs(dlead)
      }
      toadd <-
        swm[ageCAT6 == 1 &
              person_num > 1, sample(unique(pid), addN)]
      swm[pid %in% toadd, `:=` (age = age + 5, ageCAT6 = ageCAT6 + 1)]
    }
    # Update age difference table
    newpop <- swm[ageCAT6 == 1, .N]
    newlead <- swm[ageCAT6 == 2, .N]
    diff[ageCAT6 == 1, `:=`(urbansim_hh_pop = newpop, diff = newpop - umdi_hh_pop)]
    diff[ageCAT6 == 2, `:=`(urbansim_hh_pop = newlead, diff = newlead - umdi_hh_pop)]
    suppressWarnings(rm(newlag,newpop,newlead,addN,subN,da,dlag,dlead,tosub,toadd))
    gc()
    
    # ageCAT = 4
    dlag <- diff[ageCAT6 == 3, diff]
    da <- diff[ageCAT6 == 4, diff]
    dlead <- diff[ageCAT6 == 5, diff]
    if (da > 0 & dlag < 0) {
      if (abs(dlag) > da) {
        subN <- abs(da)
      }
      else if (abs(dlag) <= da) {
        subN <- abs(dlag)
      }
      tosub <-
        swm[age %in% 15:18 &
              person_num > 1, sample(unique(pid), subN)]
      swm[pid %in% tosub, `:=` (age = age - 5, ageCAT6 = ageCAT6 - 1)]
    }
    # Update age difference table
    newpop <- swm[ageCAT6 == 4, .N]
    newlag <- swm[ageCAT6 == 3, .N]
    diff[ageCAT6 == 4, `:=`(urbansim_hh_pop = newpop, diff = newpop - umdi_hh_pop)]
    diff[ageCAT6 == 3, `:=`(urbansim_hh_pop = newlag, diff = newlag - umdi_hh_pop)]
    suppressWarnings(rm(newlag,newpop,newlead,addN,subN,da,dlag,dlead,tosub,toadd))
    gc()
    
    for (i in 6:17) {
      dlag <- diff[ageCAT6 == i - 1, diff]
      da <- diff[ageCAT6 == i, diff]
      dlead <- diff[ageCAT6 == i + 1, diff]
      
      if (da > 0 & dlag < 0 & dlead < 0) {
        if ((abs(dlag) + abs(dlead)) > da) {
          subN <- abs(round(da * (dlag / (dlag + dlead))))
          addN <- abs(round(da * (dlead / (dlag + dlead))))
        }
        else if ((abs(dlag) + abs(dlead)) <= da) {
          subN <- abs(dlag)
          addN <- abs(dlead)
        }
        tosub <-
          swm[ageCAT6 == i &
                person_num > 1, sample(unique(pid), size=min(subN,lun(pid)))]
        toadd <-
          swm[ageCAT6 == i &
                person_num > 1, sample(unique(pid), size=min(addN,lun(pid)))]
        swm[pid %in% tosub, `:=` (age = age - 5, ageCAT6 = ageCAT6 - 1)]
        swm[pid %in% toadd, `:=` (age = age + 5, ageCAT6 = ageCAT6 + 1)]
      }
      if (da > 0 & dlag < 0 & dlead > 0) {
        if (abs(dlag) > da) {
          subN <- abs(da)
        }
        else if (abs(dlag) <= da) {
          subN <- abs(dlag)
        }
        tosub <-
          swm[ageCAT6 == i &
                person_num > 1, sample(unique(pid), size=min(subN,lun(pid)))]
        swm[pid %in% tosub, `:=` (age = age - 5, ageCAT6 = ageCAT6 - 1)]
      }
      # Update age difference table
      newpop <- swm[ageCAT6 == i, .N]
      newlag <- swm[ageCAT6 == i - 1, .N]
      newlead <- swm[ageCAT6 == i + 1, .N]
      diff[ageCAT6 == i - 1, `:=`(urbansim_hh_pop = newlag, diff = newlag - umdi_hh_pop)]
      diff[ageCAT6 == i, `:=`(urbansim_hh_pop = newpop, diff = newpop - umdi_hh_pop)]
      diff[ageCAT6 == i + 1, `:=`(urbansim_hh_pop = newlead, diff = newlead - umdi_hh_pop)]
      suppressWarnings(rm(newlag,newpop,newlead,addN,subN,da,dlag,dlead,tosub,toadd))
      gc()
      
    }
    
    # ageCAT = 5
    da <- diff[ageCAT6 == 5, diff]
    dlead <- diff[ageCAT6 == 6, diff]
    if (da > 0 & dlead < 0) {
      if (abs(dlead) > da) {
        addN <- abs(da)
      }
      else if (abs(dlead) <= da) {
        addN <- abs(dlead)
      }
      toadd <-
        swm[ageCAT6 == 5 &
              person_num > 1, sample(unique(pid), size=min(addN,lun(pid)))]
      swm[pid %in% toadd, `:=` (age = age + 5, ageCAT6 = ageCAT6 + 1)]
    }
    # Update age difference table
    newpop <- swm[ageCAT6 == 5, .N]
    newlead <- swm[ageCAT6 == 6, .N]
    diff[ageCAT6 == 5, `:=`(urbansim_hh_pop = newpop, diff = newpop - umdi_hh_pop)]
    diff[ageCAT6 == 6, `:=`(urbansim_hh_pop = newlead, diff = newlead - umdi_hh_pop)]
    
    suppressWarnings(rm(newlag,newpop,newlead,addN,subN,da,dlag,dlead,tosub,toadd))
    gc()
    
    # ageCAT = 18
    da <- diff[ageCAT6 == 18, diff]
    dlag <- diff[ageCAT6 == 17, diff]
    if (da > 0 & dlag < 0) {
      if (abs(dlag) > da) {
        subN <- abs(da)
      }
      else if (abs(dlag) <= da) {
        subN <- abs(dlag)
      }
      avail <-  swm[ageCAT6 == 18 &
                      person_num > 1, .N]
      tosub <-
        swm[ageCAT6 == 18 &
              person_num > 1, sample(unique(pid), min(avail, subN))]
      swm[pid %in% tosub, `:=` (age = age - 5, ageCAT6 = ageCAT6 - 1)]
    }
    if (da < 0 & dlag > 0) {
      if (abs(da) > dlag) {
        addN <- abs(dlag)
      }
      else if (abs(da) <= dlag) {
        addN <- abs(da)
      }
      toadd <-
        swm[ageCAT6 == 17 &
              person_num > 1, sample(unique(pid), size=min(addN,lun(pid)))]
      swm[pid %in% toadd, `:=` (age = age + 5, ageCAT6 = ageCAT6 + 1)]
    }
    # Update age difference table
    newpop <- swm[ageCAT6 == 18, .N]
    newlag <- swm[ageCAT6 == 17, .N]
    diff[ageCAT6 == 18, `:=`(urbansim_hh_pop = newpop, diff = newpop - umdi_hh_pop)]
    diff[ageCAT6 == 17, `:=`(urbansim_hh_pop = newlag, diff = newlag - umdi_hh_pop)]
    
    suppressWarnings(rm(newlag,newpop,newlead,addN,subN,da,dlag,dlead,tosub,toadd))
    gc()
    
    # Recalculate household characteristics
    # Persons by 5-yr age categories
    swm[age < 5 , ageCAT6 := 1]
    swm[age %in% 5:9 , ageCAT6 := 2]
    swm[age %in% 10:14 , ageCAT6 := 3]
    swm[age %in% 15:19 , ageCAT6 := 4]
    swm[age %in% 20:24 , ageCAT6 := 5]
    swm[age %in% 25:29 , ageCAT6 := 6]
    swm[age %in% 30:34 , ageCAT6 := 7]
    swm[age %in% 35:39 , ageCAT6 := 8]
    swm[age %in% 40:44 , ageCAT6 := 9]
    swm[age %in% 45:49 , ageCAT6 := 10]
    swm[age %in% 50:54 , ageCAT6 := 11]
    swm[age %in% 55:59 , ageCAT6 := 12]
    swm[age %in% 60:64 , ageCAT6 := 13]
    swm[age %in% 65:69 , ageCAT6 := 14]
    swm[age %in% 70:74 , ageCAT6 := 15]
    swm[age %in% 75:79 , ageCAT6 := 16]
    swm[age %in% 80:84 , ageCAT6 := 17]
    swm[age >= 85 , ageCAT6 := 18]
    # Count of persons by ageCAT6 in each HHd
    swm[ageCAT6 == 1, age1 := .N, hid]
    swm[ageCAT6 == 2, age2 := .N, hid]
    swm[ageCAT6 == 3, age3 := .N, hid]
    swm[ageCAT6 == 4, age4 := .N, hid]
    swm[ageCAT6 == 5, age5 := .N, hid]
    swm[ageCAT6 == 6, age6 := .N, hid]
    swm[ageCAT6 == 7, age7 := .N, hid]
    swm[ageCAT6 == 8, age8 := .N, hid]
    swm[ageCAT6 == 9, age9 := .N, hid]
    swm[ageCAT6 == 10, age10 := .N, hid]
    swm[ageCAT6 == 11, age11 := .N, hid]
    swm[ageCAT6 == 12, age12 := .N, hid]
    swm[ageCAT6 == 13, age13 := .N, hid]
    swm[ageCAT6 == 14, age14 := .N, hid]
    swm[ageCAT6 == 15, age15 := .N, hid]
    swm[ageCAT6 == 16, age16 := .N, hid]
    swm[ageCAT6 == 17, age17 := .N, hid]
    swm[ageCAT6 == 18, age18 := .N, hid]
    dt_shift <- rbind(dt_shift, swm)
    message(paste0(m, ' complete ', yr))
  }
  saveRDS(dt_shift,paste0(output_path,'pop_shifted_',scen,'_',yr,'.rds'))
}

# Recalculate age differences to UMDI
#fs <-
#  list.files(path = output_path,
#             pattern = 'pop_shifted_',
#             full.names = TRUE)

#readComp <- function(x) {
#    dt <- readRDS(x)
#    dt[, pop := 1]
#    dt <-
#      dt[, lapply(.SD, sum), .(mpo, year, ageCAT6), .SDcols = 'pop']

#  urbsim <- lapply(fs, readComp)
#  uspop <- rbindlist(urbsim)
  uspop <- dt_shift[, lapply(.SD, sum), .(mpo, year, ageCAT6), .SDcols = 'pop']
  uspop <- uspop[!is.na(mpo)]
  setkey(uspop, mpo, year, ageCAT6)
  # adiff <- umdi.swm[uspop]
  adiff <- umdi[uspop]
  adiff[, diff := pop - umdi_hh_pop]
  setnames(adiff, 'pop', 'urbansim_hh_pop')
  fwrite(adiff,paste0(output_path,'post_shift_umdi_pop_age_difference_',yr,'.csv'))

rm(dt_shift);gc()

```

# 8.3 - Person duplication function

```{r}
#| eval: false

# Function that creates subsets of households to sample from for replicate/delete adjustments
cx <- function(x) {
  sub1 <- x[new_WGTP==1 & ageCAT6 == 1  & children > 1]
  add1 <- x[new_WGTP==1 & ageCAT6 == 1  & children < 4 &
              adults > 1 & ageHHder %in% 20:60]
  sub2 <- x[new_WGTP==1 & ageCAT6 == 2  & children > 1]
  add2 <- x[new_WGTP==1 & ageCAT6 == 2  & children < 4 &
              adults > 1 & ageHHder %in% 20:60]
  sub3 <- x[new_WGTP==1 & ageCAT6 == 3  & children > 1]
  add3 <- x[new_WGTP==1 & ageCAT6 == 3  & children < 4 &
              adults > 1 & ageHHder %in% 20:60]
  sub4 <-
    x[(new_WGTP==1 & ageCAT6 == 4  &
         HHT == 3 &
         persons >= 2 & persons<8 & person_num > 1) | (new_WGTP==1 & ageCAT6 == 4 &
                            HHT == 1 & children > 1 & person_num > 1)]
  add4 <-
    x[(new_WGTP==1 & ageCAT6 == 4  &
         HHT == 1 &
         children < 4 &
         adults > 1 &
         ageHHder %in% 25:50 &
         person_num > 1) |
        (new_WGTP==1 & ageCAT6 == 4  & ageHHder < 24 & HHT == 3 & persons < 4 &
           person_num > 1)]
  
  sub5 <-
    x[(new_WGTP==1 & ageCAT6 == 5  &
         HHT == 3 &
         persons > 2  & persons<8 &
         person_num > 1) | (new_WGTP==1 & ageCAT6 == 5 &
                              HHT == 1 & adults >= 2 &
                              person_num > 1)]
  add5 <- x[ageCAT6 == 5  & HHT == 3 &
              persons >= 2  & persons<8 & ageHHder %in% 20:44]
  
  sub6 <-
    x[(new_WGTP==1 & ageCAT6 == 6  &
         HHT == 3 &
         persons > 2  & persons<8 &
         person_num > 1) |
        (new_WGTP==1 & ageCAT6 == 6  & HHT == 1 & adults >= 2 &
           person_num > 1)]
  add6 <- x[ageCAT6 == 6  & HHT == 3 &
              persons >= 2  & persons<8 & ageHHder %in% 20:44]
  
  sub7 <-
    x[(new_WGTP==1 & ageCAT6 == 7  &
         HHT == 3 &
         persons > 2  & persons<8 &
         person_num > 1) |
        (new_WGTP==1 & ageCAT6 == 7  & HHT == 1 & adults >= 2 &
           person_num > 1)]
  add7 <- x[new_WGTP==1 & ageCAT6 == 7  & HHT == 3 &
              persons >= 2  & persons<8 & ageHHder %in% 20:44]
  
  sub8 <-
    x[(new_WGTP==1 & ageCAT6 == 8  &
         HHT == 3 &
         persons > 2  & persons<8 &
         person_num > 1) |
        (new_WGTP==1 & ageCAT6 == 8  & HHT == 1 & adults >= 2 &
           person_num > 1)]
  add8 <- x[ageCAT6 == 8  & HHT == 3 &
              persons >= 2  & persons<8 & ageHHder %in% 24:49]
  
  sub9 <-
    x[(new_WGTP==1 & ageCAT6 == 9  &
         HHT == 3 &
         persons > 2  & persons<8 &
         person_num > 1) | (new_WGTP==1 & ageCAT6 == 9 &
                              HHT == 1 & adults >= 2 &
                              person_num > 1)]
  add9 <- x[new_WGTP==1 & ageCAT6 == 9  & HHT == 3 &
              persons >= 2  & persons<8 & ageHHder %in% 30:54]
  
  sub10 <-
    x[(new_WGTP==1 & ageCAT6 == 10 &
         HHT == 3 &
         persons > 2  & persons<8 &
         person_num > 1) |
        (new_WGTP==1 & ageCAT6 == 10 & HHT == 1 & adults >= 2 &
           person_num > 1)]
  add10 <- x[ageCAT6 == 10 &
               HHT == 3 & persons >= 2  & persons<8 & ageHHder %in% 40:59]
  
  sub11 <-
    x[(new_WGTP==1 & ageCAT6 == 11 &
         HHT == 3 &
         persons > 2  & persons<8 &
         person_num > 1) |
        (new_WGTP==1 & ageCAT6 == 11 &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add11 <- x[ageCAT6 == 11 &
               HHT == 3 & persons >= 2  & persons<8 & ageHHder %in% 40:64]
  sub12 <-
    x[(new_WGTP==1 & ageCAT6 == 12 &
         HHT == 3 &
         persons > 2  & persons<8 &
         person_num > 1) |
        (new_WGTP==1 & ageCAT6 == 12 &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add12 <- x[new_WGTP==1 & ageCAT6 == 12  &
               HHT == 3 & persons >= 2  & persons<8 & ageHHder %in% 45:69]
  sub13 <-
    x[(new_WGTP==1 & ageCAT6 == 13  &
         HHT == 3 &
         persons >= 2  & persons<8 &
         person_num > 1) |
        (new_WGTP==1 & ageCAT6 == 13  &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add13 <- x[ageCAT6 == 13  &
               HHT == 3 & persons >= 2 & ageHHder %in% 50:74]
  
  sub14 <-
    x[(new_WGTP==1 & ageCAT6 == 14  &
         HHT == 3 &
         persons >= 2  & persons<8 &
         person_num > 1) |
        (new_WGTP==1 & ageCAT6 == 14  &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add14 <- x[ageCAT6 == 14 &
               HHT > 1 & persons >= 2 & persons<8 ]
  sub15 <-
    x[(new_WGTP==1 & ageCAT6 == 15 &
         HHT == 3 &
         persons > 2  & persons<8 &
         person_num > 1) |
        (new_WGTP==1 & ageCAT6 == 15  &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add15 <- x[new_WGTP==1 & ageCAT6 == 15 &
               HHT > 1 & persons >= 2 & persons<8 ]
  sub16 <-
    x[(new_WGTP==1 & ageCAT6 == 16 &
         HHT == 3 &
         persons > 2  & persons<8 &
         person_num > 1) |
        (new_WGTP==1 & ageCAT6 == 16 &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add16 <- x[ageCAT6 == 16 &
               HHT > 1 & persons >= 2 & persons<8 ]
  sub17 <-
    x[(new_WGTP==1 & ageCAT6 == 17 &
         HHT == 3 &
         persons > 2  & persons<8 &
         person_num > 1) |
        (new_WGTP==1 & ageCAT6 == 17 &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add17 <- x[new_WGTP==1 & ageCAT6 == 17 &
               HHT > 1 & persons >= 2 & persons<8 ]
  sub18 <-
    x[(new_WGTP==1 & ageCAT6 == 18 &
         HHT == 3 &
         persons >= 2  & persons<8 &
         person_num > 1) |
        (new_WGTP==1 & ageCAT6 == 18 &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add18 <- x[ageCAT6 == 18 &
               HHT > 1 & persons >= 2 & persons<8 ]
  
  add.sample <<-
    list(
      add1,
      add2,
      add3,
      add4,
      add5,
      add6,
      add7,
      add8,
      add9,
      add10,
      add11,
      add12,
      add13,
      add14,
      add15,
      add16,
      add17,
      add18
    )
  sub.sample <<-
    list(
      sub1,
      sub2,
      sub3,
      sub4,
      sub5,
      sub6,
      sub7,
      sub8,
      sub9,
      sub10,
      sub11,
      sub12,
      sub13,
      sub14,
      sub15,
      sub16,
      sub17,
      sub18
    )
  
}
```

# 8.4 - Post-processor 04 - run person duplication/subtraction

```{r}

flaglist <- data.table()


yr=2035
# Load shift-corrected population differences
# adiff <-
  # fread(paste0(output_path,'post_shift_umdi_pop_age_difference_',yr,'.csv'))

adiff <- fread(paste0(output_path,'umdi_pop_age_difference_',yr,'.csv'))
adiff[,diff:=round(diff)]


  dt <- fread(paste0(output_path,'expanded_reweighted_PUMS_',yr,'.csv'))
#  dt <-
 #   readRDS(paste0(output_path,'pop_shifted_',scen,'_',yr,'.rds'))
  dt[ageCAT6 == 1, age1 := .N, hid]
  dt[ageCAT6 == 2, age2 := .N, hid]
  dt[ageCAT6 == 3, age3 := .N, hid]
  dt[ageCAT6 == 4, age4 := .N, hid]
  dt[ageCAT6 == 5, age5 := .N, hid]
  dt[ageCAT6 == 6, age6 := .N, hid]
  dt[ageCAT6 == 7, age7 := .N, hid]
  dt[ageCAT6 == 8, age8 := .N, hid]
  dt[ageCAT6 == 9, age9 := .N, hid]
  dt[ageCAT6 == 10, age10 := .N, hid]
  dt[ageCAT6 == 11, age11 := .N, hid]
  dt[ageCAT6 == 12, age12 := .N, hid]
  dt[ageCAT6 == 13, age13 := .N, hid]
  dt[ageCAT6 == 14, age14 := .N, hid]
  dt[ageCAT6 == 15, age15 := .N, hid]
  dt[ageCAT6 == 16, age16 := .N, hid]
  dt[ageCAT6 == 17, age17 := .N, hid]
  dt[ageCAT6 == 18, age18 := .N, hid]
  dt[, sporder := person_num]
  dt[, children := sum(child), hid]
  dt[child == 0, adult := 1]
  dt[, adults := sum(adult), hid]
  dt[SPORDER == 1, ageHH := AGEP]
  dt[, ageHHder := max(ageHH, na.rm = T), hid]
  dt[, persons := max(SPORDER), hid]
  dt[SPORDER == 1, ht := HHtype]
  dt[, HHT := max(ht, na.rm = T), hid]
  addon <- data.table()
  delflag <- data.table()
  
  for (m in mpos) {
    swm <- dt[mpo == m]
    cx(swm)
    
    for (i in 1:3) {
      diff <- adiff[mpo == m & ageCAT6 == i & year == yr, diff]
      if (diff < 0) {
        sN <- abs(diff)
        m1 <- add.sample[[i]]
        if (lun(m1$pid) < sN) {
          m1 <- swm[ageCAT6 == i  & children < 4 &
                       adults >= 1 & ageHHder %in% 20:60]
        }
        if (lun(m1$pid) < sN) {
          m1 <- swm[ageCAT6 == i  & children <= 4]
        }
        if (lun(m1$pid) < sN) {
          m1 <- swm[ageCAT6 == i]
        }
        hhs <- sample(unique(m1$hid), min(lun(m1$hid), sN))
        if (lun(m1$hid) < sN) {
          flaglist <-
            c(flaglist, data.table(
              flag = paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i
              )
            ))
          message(
            paste0(
              'WARNING - Insufficient sample size for duplication ',
              m,
              ' ',
              yr,
              ' - ',
              'ageCAT == ',
              i
            )
          )
        }
        samp <- m1[hid %in% hhs & ageCAT6 == i]
        samp1 <- samp[get(paste0('age', i)) == 1]
        samp2 <-
          samp[get(paste0('age', i)) > 1, .SD[sample(.N, 1)], hid]
        addon <- rbind(addon, samp1, samp2)
      }
      if (diff > 0) {
        sN <- abs(diff)
        m1 <- sub.sample[[i]]
        hhs <- sample(unique(m1$pid),  min(lun(m1$pid), sN))
        if (lun(m1$pid) < sN) {
          flaglist <-
            c(flaglist, data.table(
              flag = paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            ))
          message(
            paste0(
              'WARNING - Insufficient sample size for duplication ',
              m,
              ' ',
              yr,
              ' - ',
              'ageCAT == ',
              i,
              ' --- diff = ',
              diff,
              ' --- sample = ',
              dim(m1)[1]
            )
          )
        }
        samp <- m1[pid %in% hhs]
        samp[, delete := 1]
        delflag <- rbind(delflag, samp[, .(pid, delete)])
      }
    }
    
    for (i in 4) {
      diff <- adiff[mpo == m & ageCAT6 == i & year == yr, diff]
      if (diff < 0) {
        sN <- abs(diff)
        m1 <- add.sample[[i]]
        hhs <- sample(unique(m1$hid), min(lun(m1$hid), sN))
        if (lun(m1$hid) < sN) {
          flaglist <-
            c(flaglist, data.table(
              flag = paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i
              )
            ))
          message(
            paste0(
              'WARNING - Insufficient sample size for duplication ',
              m,
              ' ',
              yr,
              ' - ',
              'ageCAT == ',
              i
            )
          )
        }
        samp1 <-
          m1[hid %in% hhs &
               HHT == 1 & ageCAT6 == i, .SD[sample(.N, 1)], hid]
        samp1[, age := sample(15:18, 1)]
        samp2 <-
          m1[hid %in% hhs &
               HHT == 3 &
               ageCAT6 == i, .SD[sample(.N, 1)], hid]
        samp2[, age := 19][, ageCAT6 := 4]
        addon <- rbind(addon, samp1, samp2)
      }
      if (diff > 0) {
        sN <- abs(diff)
        m1 <- sub.sample[[i]]
        hhs <- sample(unique(m1$pid),  min(lun(m1$pid), sN))
        if (lun(m1$pid) < sN) {
          flaglist <-
            c(flaglist, data.table(
              flag = paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            ))
          message(
            paste0(
              'WARNING - Insufficient sample size for duplication ',
              m,
              ' ',
              yr,
              ' - ',
              'ageCAT == ',
              i,
              ' --- diff = ',
              diff,
              ' --- sample = ',
              dim(m1)[1]
            )
          )
        }
        samp <- m1[pid %in% hhs]
        samp[, delete := 1]
        delflag <- rbind(delflag, samp[, .(pid, delete)])
      }
    }
    for (i in 5:10) {
      diff <- adiff[mpo == m & ageCAT6 == i & year == yr, diff]
      if (diff < 0) {
        sN <- abs(diff)
        if (m %in% c('CCC',
                     'FRCOG',
                     'MRPC',
                     'MVC',
                     'NPEDC',
                     'NMCOG',
                     'OCPC',
                     'PVPC')) {
          m1 <- swm[ageCAT6 == i & HHT == 3]
          if (lun(m1$pid) < sN) {
              m1 <- swm[ageCAT6 == i & HHT < 5]
            }
          hhs <- sample(unique(m1$pid), min(lun(m1$pid), sN))
          samp <- m1[pid %in% hhs]
          if (lun(m1$pid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for duplication ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            )
          }
        }
        if (!m %in% c('CCC',
                      'FRCOG',
                      'MRPC',
                      'MVC',
                      'NPEDC',
                      'NMCOG',
                      'OCPC',
                      'PVPC')) {
          m1 <- add.sample[[i]]
          hhs <- sample(unique(m1$pid), min(lun(m1$pid), sN))
          if (lun(m1$pid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for duplication ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            )
          }
          samp <- m1[pid %in% hhs]
        }
        addon <- rbind(addon, samp)
      }
        if (diff > 0) {
          sN <- abs(diff)
          m1 <- sub.sample[[i]]
          hhs <- sample(unique(m1$pid),  min(lun(m1$pid), sN))
          if (lun(m1$pid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for duplication ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            )
          }
          samp <- m1[pid %in% hhs]
          samp[, delete := 1]
          delflag <- rbind(delflag, samp[, .(pid, delete)])
        }
      }
      
      
      for (i in 11:18) {
        diff <- adiff[mpo == m & ageCAT6 == i & year == yr, diff]
        if (diff < 0) {
          sN <- abs(diff)
          m1 <- add.sample[[i]]
          hhs <- sample(unique(m1$pid), min(lun(m1$pid), sN))
          if (lun(m1$pid) < sN) {
            m1 <- swm[ageCAT6 == i & HHT == 3]
          }
          if (lun(m1$pid) < sN) {
            m1 <- swm[ageCAT6 == i & HHT < 5]
          }
          if (lun(m1$pid) < sN) {
            m1 <- swm[ageCAT6 == i]
          }
          if (lun(m1$pid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for duplication ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            )
          }
          samp <- m1[pid %in% hhs]
          addon <- rbind(addon, samp)
        }
        if (diff > 0) {
          sN <- abs(diff)
          m1 <- sub.sample[[i]]
          if (lun(m1$pid) < sN){
            m1 <- swm[ageCAT6 == i & person_num > 1]
          }
          hhs <- sample(unique(m1$pid),  min(lun(m1$pid), sN))
          if (lun(m1$pid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for duplication ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            )
          }
          samp <- m1[pid %in% hhs]
          samp[, delete := 1]
          delflag <- rbind(delflag, samp[, .(pid, delete)])
        }
      }
        message(paste0(m, ' complete ', yr))
        suppressWarnings(rm(add.sample,sub.sample))
    }
  
    # Join IDs of persons to delete back to main data.table and then delete those rows
    setkey(delflag, pid)
    setkey(dt, pid)
    dt2 <- delflag[dt]
    dt2 <- dt2[is.na(delete)]
    
    # Join person records to append to data.table
    addon[, addflag := 1]
    dt3 <- rbind(dt2, addon, fill = T)
    
    # Recalculate person and household characteristics
    setorder(dt3, hid, person_num)
    dt3[, person_num := 1:.N, hid]
    dt3[, sporder := person_num]
    dt3[, pid := paste0(hid, person_num)]
    dt3[, persons := .N, hid]
    dt3[, workers := sum(worker), hid]
    dt3[, pinc2021:= as.numeric(ADJINC) * PINCP]
    dt3[, HINCP.adj.2021USD := sum(pinc2021), hid]
    dt3[, HHsize:=max(sporder),hid]
    
    
    dt3 <- dt3 %>% mutate(
      child = case_when(age < 18 ~ 1,
                        age >= 18 ~ 0)) |>  
      setDT()
    # Count children and adults in each household
    dt3[, children_in_HH := sum(child, na.rm = T), hid]
    dt3[child == 0, adult := 1][is.na(adult), adult := 0]
    dt3[, adults := sum(adult, na.rm = T), hid]

    # Persons by 5-yr age categories
    dt3[age < 5 , ageCAT6 := 1]
    dt3[age %in% 5:9 , ageCAT6 := 2]
    dt3[age %in% 10:14 , ageCAT6 := 3]
    dt3[age %in% 15:19 , ageCAT6 := 4]
    dt3[age %in% 20:24 , ageCAT6 := 5]
    dt3[age %in% 25:29 , ageCAT6 := 6]
    dt3[age %in% 30:34 , ageCAT6 := 7]
    dt3[age %in% 35:39 , ageCAT6 := 8]
    dt3[age %in% 40:44 , ageCAT6 := 9]
    dt3[age %in% 45:49 , ageCAT6 := 10]
    dt3[age %in% 50:54 , ageCAT6 := 11]
    dt3[age %in% 55:59 , ageCAT6 := 12]
    dt3[age %in% 60:64 , ageCAT6 := 13]
    dt3[age %in% 65:69 , ageCAT6 := 14]
    dt3[age %in% 70:74 , ageCAT6 := 15]
    dt3[age %in% 75:79 , ageCAT6 := 16]
    dt3[age %in% 80:84 , ageCAT6 := 17]
    dt3[age >= 85 , ageCAT6 := 18]
    
    cf <- dt3[,lapply(.SD,sum),ageCAT6,.SDcols='new_WGTP']
    setnames(cf,'new_WGTP','pop_in_HH')
    cf[,year:=2035][,pop_in_HH:=round(pop_in_HH)]
    setkey(cf,ageCAT6)
    ad <- adiff[year==2035,lapply(.SD,sum),ageCAT6,.SDcols = c('umdi_hh_pop','urbansim_hh_pop')]
    setkey(ad,ageCAT6)
    
    cc <- ad[cf]
    cc[,diff:=pop_in_HH-umdi_hh_pop]
    
    ct <- dt3[,lapply(.SD,sum),.(ageCAT6,HHT),.SDcols='new_WGTP']
    setnames(ct,'new_WGTP','pop_in_HH')
    ct[,year:=2035][,pop_in_HH:=round(pop_in_HH)]
    setorder(ct,ageCAT6,HHT)
    
    saveRDS(
      dt3,
      paste0(output_path,
        'pop_corrected_PUMS_',scen,'_',yr,'.csv'
      )
    )
    }
  
```
