---
title: "03_Reweighter.Output.Summaries"
format: html
editor: visual
---

# 0.0 Environment Set Up

# 0.1 Filepath Set Up

```{r}
#===============================================================================
#TODO: 
# 1. Create a way to loop through scenarios and headship rates based on a list

#===============================================================================
### Set Scenario Names ###
#Population Scenario
scen <- 'Scenario_1'
#Headship rate Scenario
hdshp <- 'acs1721'

#Use when Rproject is open
#root <- '../../../../Data/Working/Regional_Control_Totals/'

# Base root. Toggle between S: and K: drive depedning on mapped paths on your computer
base <- "K:/DataServices/Projects/Current_Projects/"
#base <- "S:/Network Shares/K Drive/DataServices/Projects/Current_Projects/"

# General K drive filepath
root <- paste0(base, "Housing/StatewideHousingPlan/04_Analysis/Data/Working/Regional_Control_Totals/")

# Set output filepath
output_path <- paste0(base, "Housing/StatewideHousingPlan/04_Analysis/Data/Final_Outputs/", scen, "/")

# Reweighter files path
rwt_path <- paste0(base, "Housing/StatewideHousingPlan/04_Analysis/Data/Working/Reweighter/", scen, "/Output_Files/")

# Set knitr path
knitr::opts_knit$set(root.dir = root)
```

# 0.2 Set Package Dependencies + Set Environment Variables, Set up data keys

```{r}
library(tidycensus)
library(tidyverse)
library(data.table)
library(reticulate)
library(mapcdatakeys)
library(janitor)
library(readxl)

##get rid of scientific notation
options(scipen = 999)
#Set random seed
set.seed(351)

#Set up muni data keys
#MPO Data Keys
keys.mpo <- mapcdatakeys::all_muni_data_keys |> 
  select(
    muni_id, muni_name, mpo
  ) |> 
  mutate(
    muni_id = as.character(muni_id),
    muni_name = ifelse(muni_name == "Manchester-by-the-Sea", "Manchester", muni_name)
  )

#RPA Data Keys
keys.rpa <- mapcdatakeys::all_muni_data_keys |> 
  select(
    muni_id, muni_name, rpa_acr
  ) |> 
  mutate(
    muni_id = as.character(muni_id),
    muni_name = ifelse(muni_name == "Manchester-by-the-Sea", "Manchester", muni_name)
  )
```

# 1.0 Load Reweighter Output Data

```{r}
# Set a column of years corresponding to the projection years considered
years <- c(2020,2030,2035,2040,2045,2050)

# Create null objects for each scenario's full reweighted PUMS dataset. 
# One will be used per script run depending on the value given to "scen" at the outset of the script.
s1.df <- NULL

s2.sf <- NULL

s3.sf <- NULL

s4.df <- NULL

# Standardized function for reading reweighted PUMS data for each scenario and projection year.
# - scenario: data that is generated to reflect population and headship rate scenarios. See xx for further detail
# - year: a year in a list of years created earlier in this code block. Typically in 5-year or 10-year intervals.
read_rwt.files <- function(scenario, year){
  read.csv(
      paste0(rwt_path, "reweighted_PUMS_",scen,"_MAPC101_",i,".csv")
    )
}

# Loads in reweighted PUMS data for each year and scenario.
# Compiles data into a single dataframe which is generated earlier.
for (i in years) {
  for (j in scen){
    
    if (scen == "Scenario_1"){
    
      # Reads in the reweighted PUMS data by year.
      tmp <- read_rwt.files(scenario = scen, year = i)
    
      # Appends to dataframe.
      s1.df <- bind_rows(s1.df, tmp)
    
      # Removes the temporary file.
      rm(tmp)
    
      # Print statement when loop has completed running.
      print(paste0("Finished Reading and Appending Scenario: ", scen, " for Year: ", i))
  
  }
    else if (scen == "Scenario_2"){
    
      # Reads in the reweighted PUMS data by year.
      tmp <- read_rwt.files(scenario = scen, year = i)
    
      # Appends to dataframe.
      s2.df <- bind_rows(s2.df, tmp)
    
      # Removes the temporary file.
      rm(tmp)
    
      # Print statement when loop has completed running.
      print(paste0("Finished Reading and Appending Scenario: ", scen, " for Year: ", i))
    
  }
    else if (scen == "Scenario_3"){
    
      # Reads in the reweighted PUMS data by year.
      tmp <- read_rwt.files(scenario = scen, year = i)
    
      # Appends to dataframe.
      s3.df <- bind_rows(s3.df, tmp)
    
      # Removes the temporary file.
      rm(tmp)
    
      # Print statement when loop has completed running.
      print(paste0("Finished Reading and Appending Scenario: ", scen, " for Year: ", i))
    
  }
   else{
    
      # Reads in the reweighted PUMS data by year.
      tmp <- read_rwt.files(scenario = scen, year = i)
    
      # Appends to dataframe.
      s4.df <- bind_rows(s4.df, tmp)
    
      # Removes the temporary file.
      rm(tmp)
    
      # Print statement when loop has completed running.
      print(paste0("Finished Reading and Appending Scenario: ", scen, " for Year: ", i))
    
  }
 }
}

```

# 2.0 Output Tables For UrbanSim

-   .1 UrbanSim Household Income Groupings
-   .2 5-year Age Groupings (ageCAT6)
-   .3 Household Type

```{r}
#===============================================================================
# .1 UrbanSim Household Income Groupings
tbl.1 <- s1.df |> 
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(
    rpa_acr,
    year,
    WAGPALLC_US
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    Income_Group = case_when(
      WAGPALLC_US == 1 ~ "<=$35,000",
      WAGPALLC_US == 2 ~ "$35,001 - $75,000",
      WAGPALLC_US == 3 ~ "$75,001 - $125,000",
      WAGPALLC_US == 4 ~ "$125,001 - $225,000",
      WAGPALLC_US == 5 ~ ">$225,000",
      .default = "Unknown"
    )
  ) |> 
  select(
    -c(WAGPALLC_US)
  ) |> relocate(
    Income_Group,
    .before = hhds
  )

#Write Output tables to K Drive
write.csv(
  tbl.1,
  paste0(output_path,"/output.tbl_Households.by.IncomeGroups.csv"),
  row.names = FALSE
)

#===============================================================================
# .2 5-year Age Groupings (ageCAT6)
tbl.2 <- s1.df |> 
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(
    rpa_acr,
    year,
    ageCAT6
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    Age_Group = case_when(
      ageCAT6 == 1 ~ "00-04",
      ageCAT6 == 2 ~ "05-09",
      ageCAT6 == 3 ~ "10-14",
      ageCAT6 == 4 ~ "15-19",
      ageCAT6 == 5 ~ "20-24",
      ageCAT6 == 6 ~ "25-29",
      ageCAT6 == 7 ~ "30-34",
      ageCAT6 == 8 ~ "35-39",
      ageCAT6 == 9 ~ "40-44",
      ageCAT6 == 10 ~ "45-49",
      ageCAT6 == 11 ~ "50-54",
      ageCAT6 == 12 ~ "55-59",
      ageCAT6 == 13 ~ "60-64",
      ageCAT6 == 14 ~ "65-69",
      ageCAT6 == 15 ~ "70-74",
      ageCAT6 == 16 ~ "75-79",
      ageCAT6 == 17 ~ "80-84",
      ageCAT6 == 18 ~ "85 plus",
      .default = "Unknown"
    )
  ) |> 
  select(
    -c(ageCAT6)
  ) |> relocate(
    Age_Group,
    .before = hhds
  )

#Write Output Tables ot K Drive
write.csv(
  tbl.2,
  paste0(output_path,"/output.tbl_Households.by.AgeGroups.csv"),
  row.names = FALSE
)

#===============================================================================
# .3 Household Type
tbl.3 <- s1.df |> 
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(
    rpa_acr,
    year,
    HHtype
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    HHtype_Group = case_when(
      HHtype == 1 ~ "Household w Children",
      HHtype == 3 ~ "Household w 2 plus Adults wo Children",
      HHtype == 5 ~ "Single Person Household",
      .default = "Unknown"
    )
  ) |> 
  select(
    -c(HHtype)
  ) |> relocate(
    HHtype_Group,
    .before = hhds
  )

# Write output tables to K drive
write.csv(
  tbl.3,
  paste0(output_path,"/output.tbl_Households.by.HHtypeGroups.csv"),
  row.names = FALSE
)

```

#3.0 Output Tables for Statewide Housing Plan Deliverables - .1 Number of Children in a Household - .2 Size of Household - .3 Tenure - .4 Number of Workers in a Household

```{r}
#===============================================================================
# .1 Number of Children in a Household

#===============================================================================
# .2 Size of Household

#===============================================================================
# .3 Tenure

#===============================================================================
# .4 Number of Workers in a Household
```
