---
title: "05_Multi.Adult.Households"
format: html
editor: visual
---

# 0.0 Set-up, Define helper functions, set knitr path, load necessary packages, set muni data keys

```{r}
library(tidycensus)
library(tidyverse)
library(survey)
library(srvyr)
library(data.table)
library(janitor)
library(dplyr)
library(knitr)
library(readxl)
library(writexl)
library(openxlsx)
library(devtools)

set.seed(351)

options(scipen = 999)
```

You can add options to executable code like this

```{r, setup, include=FALSE}
#Use when Rproject is open
#root <- '../../../../Data/Working/Regional_Control_Totals/'

#General K drive filepath
root <-"K:/DataServices/Projects/Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Working/Regional_Control_Totals/"

#PUMS K drive filepath
pums_path <- "K:/DataServices/Projects/Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Working/PUMS/ipums_2021/"

rpa_path <- "K:/DataServices/Projects/Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Working/Reweighter/PUMS_data/"

#Set output filepath
output_path <- "K:/DataServices/Projects/Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Working/PUMS/outputs/"

#Set knitr path
knitr::opts_knit$set(root.dir = pums_path)
```

# 0.2 - Set package dependencies, set environment, set up data keys\*\*

```{r}
pums_raw <-read_csv("K:/DataServices/Projects/Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Working/PUMS/ipums_2021/pums_multiadult_2021_V_2024-11-22.csv")

test <- pums_raw |>
  filter(SPORDER %in% c(0,1)) |>
  group_by(SERIALNO) |>
  mutate(
    weight.test = sum(WGTP)/n()
  ) |>
  ungroup() |>
  mutate(
    weight.check = WGTP - weight.test,
    weight.check.flag = case_when(
      weight.check > 0 ~ 1,
      weight.check < 0 ~ 1,
      weight.check == 0 ~ 0,
      .default = 0
    )
  )
```

The `echo: false` option disables the printing of code (only output is displayed).

# 0.3 - PUMA to RPA Crosswalk

```{r}

# Load PUMA to RPA Crosswalk
xw <- read_csv(paste0(rpa_path, 'PUMA10_RPA_crosswalk.csv')) |> 
  mutate(
    # Change the data type and pad the string with a leading zero.
    PUMA = as.character(PUMA),
    PUMA = str_pad(PUMA, 5, side = "left", pad = "0")
  )

```

# 0.4 - PUMS Data Cleaning

```{r}
# General Data Cleaning
pums <- pums_raw |> 
  # Create a household flag for whether there is a child in a household.
  mutate( 
    Adult = case_when(
      AGEP > 18 ~ 1,
      AGEP <= 18 ~ 0,
      .default = 0
    ),
    # Create a grouping variable for types of relationships to the householder.
    RELSHIPP_Child = case_when(
      # 1 indicates a person in a household that is related to the householder by being a biological, adopted, step child,foster child, or son/daughter in law.
      RELSHIPP %in% c(25,26,27,32,35) ~ 1,
      # 2 indicates a person in a household is related to the householder outside of being a child.
      RELSHIPP %in% c(21,22,23,24,28,29,30,31,33,34,36,37,38) ~ 0,
      # Set a default flag for householders.
      .default = 0
    ),
    #  Create a grouping variable for types of relationships to the householder
    RELSHIPP_PAR = case_when(
    #1 indicates a person in a household that is a parent or parent in law.
    RELSHIPP %in% c(29,31) ~ 1,
    #2 indicates a person in a household is related to the householder outside of being a parent
    RELSHIPP %in% c(21,22,23,24,25,26,27,28,30,32,33,34,36,37,38) ~ 0,
    #Set a default flag for householders.
    .default = 0
    ),
     HINCP_adj = HINCP*ADJINC,
    # Household Income Groupings consistent with UrbanSim groupings
    HH.Income.Grp = case_when(
      HINCP_adj <= 35000 ~ "<=$35,000",
      HINCP_adj > 35000 & HINCP_adj <= 75000 ~ "$35,001-$75,000",
      HINCP_adj > 75000 & HINCP_adj <= 125000 ~ "$75,001-$125,000",
      HINCP_adj > 125000 & HINCP_adj <= 225000 ~ "$125,001-$225,000",
      HINCP_adj > 225000 ~ ">$225,000",
      .default = "0"
    ),
      # Householder Race Groupings. 
    HHLDER.RACE.Grp = case_when(
      HHLDRRAC1P == "1" ~ "White Alone",
      HHLDRRAC1P == "2" ~ "Black or African American Alone",
      HHLDRRAC1P %in% c("3", "4", "5") ~ "AIAN Alone",
      HHLDRRAC1P == "6" ~ "Asian Alone",
      HHLDRRAC1P == "7" ~ "NHPI Alone",
      HHLDRRAC1P == "8" ~ "SOR Alone",
      HHLDRRAC1P == "9" ~ "Two or More Races Alone",
      .default = "Unknown"
    ),
    # Householder Latine/hispanic Groupings
    HHLDER.HISP.Grp = case_when(
      HHLDRHISP == "01" ~ "Non-Hispanic",
      HHLDRHISP != "01" ~ "Hispanic",
      .default = "Unknown"
    ),
    # Replace any NAs (for households that own and so dont have a GRPIP and vice versa)
    GRPIP = replace_na(GRPIP, 0),
    OCPIP = replace_na(OCPIP, 0),
    # Define a "cost burdened" indicator.
    # (1) A household paying more than 30% of its income in housing costs (GRPIP or OCPIP) is considered "cost burdened"
    # (0) A household paying less than or equal to 30% of its income in housing costs is NOT considered "cost burdened"
    CostBurdened = case_when(
      GRPIP > 30 | OCPIP > 30 ~ "1",
      GRPIP <= 30 | OCPIP <= 30 ~ "0",
      .default = "0"
    ),
  # Create a variable indicating whether a person is an adult child in the household.
    AdultChild = case_when(
      Adult == 1 & RELSHIPP_Child == 1 ~ 1,
      .default = 0
    )
  )|>
  # Creating household flags for person variables.
  group_by(SERIALNO) |> 
  mutate(
    # Create a variable which is the sum of the number of children above 18
    Adult_count = sum(Adult),
    # Create a variable which is the sum of the number of son & daughter relatives in the household
    Child_count = sum(RELSHIPP_Child),
    #  Create a variable which is the sum of the number of parent relatives in the household
    Parent_count = sum(RELSHIPP_PAR),
    # Create a variable that is a count of the number of adult children in a household
    AdultChild_count = sum(AdultChild) 
  )|> 
  ungroup() |> 
  mutate(
      # Create a variable that is a household flag for whether or not a household has a adult children over 18.
    AdultChild = case_when(
      AdultChild_count > 0 ~ 1,
      Adult_count == 0 ~ 0,
      .default = 0
    ),
     # Create a variable that is a household flag for whether or not a household has a parent of the householder living in the unit.
    ParentPresent = case_when(
      Parent_count > 0 ~ 1,
      Parent_count == 0 ~ 0,
      .default = 0
    )
  )

```

# 0.5 - PUMS Analysis - State Level

The number of households with adult children

```{r}
# Number of Households with Adult Children by State-Level
householdata_adultchild <- pums |>
  filter(TYPEHUGQ == 1 & SPORDER == 1) |>
  group_by(AdultChild) |>
  summarise(hh_est = sum(WGTP)) |>
  ungroup() |>
  #Adding a column to show the total amount of households and percentage
  mutate(
    totalcount = sum(hh_est),
    percent = hh_est / totalcount
  )

#Write the output to csv
write.csv(
  householdata_adultchild,
  paste0(output_path,"05_MultiAdultHouseholds/Adult.Children.StateLevel.csv"),
  row.names = FALSE
)

# Number of Households with Adult Parents Living with Adult Children by State-Level 
householdata_parentpresent <- pums |>
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(ParentPresent) |> 
  summarise(hh_est = sum(WGTP)) |>
  ungroup() |> 
#Adding a column to show the total amount of households and percentage
   mutate(
    totalcount = sum(hh_est),
    percent = hh_est/totalcount
  )

#Write the output to csv
write.csv(
  householdata_parentpresent,
  paste0(output_path,"05_MultiAdultHouseholds/Parents.Present.StateLevel.csv"),
  row.names = FALSE
)
  
#Number of Households with Adult Children by State Level filtered by Income 
householdata_inc <- pums |>
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(AdultChild, HH.Income.Grp) |> 
  summarise(hh_est = sum(WGTP)) |> 
  ungroup() |> 
#Adding a column to show the total amount of households and percentage
  mutate(
    totalcount = sum(hh_est),
    percent = hh_est/totalcount
  )
   
#Write the output to csv
write.csv(
  householdata_inc,
  paste0(output_path,"05_MultiAdultHouseholds/Adult.Children.Inc.StateLevel.csv"),
  row.names = FALSE
)

#Number of Households with Adult Children by State Level filtered by Cost Burden
householdata_costburden <- pums |>
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(AdultChild, CostBurdened) |> 
  summarise(hh_est = sum(WGTP)) |> 
  ungroup() |> 
#Adding a column to show the total amount of households and percentage
  mutate(
    totalcount = sum(hh_est),
    percent = hh_est/totalcount
  )
   
#Write the output to csv
write.csv(
  householdata_costburden,
  paste0(output_path,"05_MultiAdultHouseholds/Adult.Children.Costburden.StateLevel.csv"),
  row.names = FALSE
)
  
#Number of Households with Adult Children by State Level filtered by Race
householdata_race <- pums |>
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(AdultChild, HHLDER.RACE.Grp, HHLDER.HISP.Grp) |> 
  summarise(hh_est = sum(WGTP)) |> 
  ungroup() |> 
#Adding a column to show the total amount of households and percentage
  mutate(
    totalcount = sum(hh_est),
    percent = hh_est/totalcount
  )

#Write the output to csv
 write.csv(
  householdata_race,
  paste0(output_path,"05_MultiAdultHouseholds/Adult.Children.Inc.Race.StateLevel.csv"),
  row.names = FALSE
)
  
```

# 0.6 - PUMS Analysis - RPA Level

```{r}
#Joining the PUMA to RPA crosswalk to the PUMS data.
pums_rpa <- left_join(
  pums,
  xw,
  by = c("PUMA")
)

hh_rpa <- pums_rpa |>
  filter(TYPEHUGQ == 1 & SPORDER == 1) |>
  group_by(rpa_acr) |> 
  summarise(hh_est = sum(WGTP)) |>
  ungroup()  

# Number of Households with Adult Children by RPA
householdata_rpa_adultchild <- pums_rpa |>
  filter(TYPEHUGQ == 1 & SPORDER == 1) |>
  group_by(rpa_acr, AdultChild) |> 
  summarise(hh_est = sum(WGTP)) |>
  ungroup() |> 
  group_by(rpa_acr) |> 
#Adding a column to show the total amount of households and percentage
  mutate(
    totalcount = sum(hh_est),
  ) |> 
  ungroup() |> 
  mutate(
    percent = hh_est/totalcount
  )

#Write the output to csv
write.csv(
  householdata_rpa_adultchild,
  paste0(output_path,"05_MultiAdultHouseholds/Adult.Children.RPA.csv"),
  row.names = FALSE
)

# Number of Households with Adult Parents Living with Adult Children by RPA
householdata_ParentPresent_rpa <- pums_rpa |>
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(rpa_acr, ParentPresent) |>
  summarise(hh_est = sum(WGTP)) |>
  ungroup()  |> 
  group_by(rpa_acr) |> 
#Adding a column to show the total amount of households and percentage
  mutate(
    totalcount = sum(hh_est),
  ) |> 
  ungroup() |> 
  mutate(
    percent = hh_est/totalcount
  )

#Write the output to csv
write.csv(
  householdata_ParentPresent_rpa,
  paste0(output_path,"05_MultiAdultHouseholds/Parents.Present.StateLevel.csv"),
  row.names = FALSE
)

#Number of Households with Adult Children by RPA filtered by Income 
householdata_rpa_inc <- pums_rpa |>
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(rpa_acr, AdultChild, HH.Income.Grp) |> 
  summarise(hh_est = sum(WGTP)) |> 
  ungroup() |> 
  group_by(rpa_acr) |> 
#Adding a column to show the total amount of households and percentage
  mutate(
    totalcount = sum(hh_est),
  ) |> 
  ungroup() |> 
  mutate(
    percent = hh_est/totalcount
  )

#Write the output to csv
write.csv(
  householdata_rpa_inc,
  paste0(output_path,"05_MultiAdultHouseholds/Adult.Children.Inc.RPA.csv"),
  row.names = FALSE
)

#Number of Households with Adult Children by RPA filtered by Race
householdata_rpa_race <- pums_rpa |>
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(rpa_acr, AdultChild, HHLDER.RACE.Grp, HHLDER.HISP.Grp) |> 
  summarise(hh_est = sum(WGTP)) |> 
  ungroup() |> 
  group_by(rpa_acr) |> 
#Adding a column to show the total amount of households and percentage
  mutate(
    totalcount = sum(hh_est),
  ) |> 
  ungroup() |> 
  mutate(
    percent = hh_est/totalcount
  )
   
#Write the output to csv
write.csv(
  householdata_rpa_race,
  paste0(output_path,"05_MultiAdultHouseholds/Adult.Children.Inc.Race.RPA.csv"),
  row.names = FALSE
)

#Number of Households with Adult Children by RPA filtered by Cost Burden
householdata_rpa_costburden <- pums_rpa |>
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(rpa_acr, AdultChild, CostBurdened) |> 
  summarise(hh_est = sum(WGTP)) |> 
  ungroup() |> 
  group_by(rpa_acr) |> 
#Adding a column to show the total amount of households and percentage
  mutate(
    totalcount = sum(hh_est),
  ) |> 
  ungroup() |> 
  mutate(
    percent = hh_est/totalcount
  )
   
#Write the output to csv
write.csv(
  householdata_rpa_costburden,
  paste0(output_path,"05_MultiAdultHouseholds/Adult.Children.Inc.Costburden.RPA.csv"),
  row.names = FALSE
)
 
  
```
